---
title: "BaitBench Probe Assessment Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  build_stats_file: ""
  build_params_file: ""
  xreact_hits_file: ""
  xreact_summary_file: ""
  xreact_threshold: 80.0
  coverage_summary_file: ""
  coverage_depth_file: ""
  coverage_multi_mapping_file: ""
  coverage_proximity: 50
  params_file: ""
  individual_coverage_file: ""
  refine_summary_file: ""
  gap_details_file: ""
  gap_min_length: 120
---

<style>
.main-container { max-width: 100%; margin-left: 1em; margin-right: 1em; }
.dataTables_wrapper { width: 100%; }
table { width: 100%; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 5, out.width = "100%")

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  library(DT)
  library(plotly)
})

theme_set(theme_minimal(base_size = 14))

# --- Load build pipeline data (optional) ---
has_build_stats <- nchar(params$build_stats_file) > 0 && file.exists(params$build_stats_file)
if (has_build_stats) {
  build_stats <- read.delim(params$build_stats_file, stringsAsFactors = FALSE)
  build_stats$step <- factor(build_stats$step, levels = build_stats$step)
  build_stats$dropped_seqs <- c(NA, -diff(build_stats$sequences))
  build_stats$dropped_bases <- c(NA, -diff(build_stats$bases))
  build_stats$pct_retained <- c(NA, round(100 * build_stats$sequences[-1] /
    build_stats$sequences[-nrow(build_stats)], 1))
}

has_build_params <- nchar(params$build_params_file) > 0 && file.exists(params$build_params_file)
if (has_build_params) {
  build_params_df <- read.delim(params$build_params_file, stringsAsFactors = FALSE)
}

# --- Load cross-reactivity data ---
hits <- read.delim(params$xreact_hits_file, stringsAsFactors = FALSE)
xreact_summary_df <- read.delim(params$xreact_summary_file, stringsAsFactors = FALSE)

threshold <- params$xreact_threshold

has_self <- any(hits$mode == "self") || any(xreact_summary_df$mode == "self")
has_against <- any(hits$mode == "against") || any(xreact_summary_df$mode == "against")

hits_self <- hits %>% filter(mode == "self")
hits_against <- hits %>% filter(mode == "against")
summary_self <- xreact_summary_df %>% filter(mode == "self")
summary_against <- xreact_summary_df %>% filter(mode == "against")

n_probes_xreact <- length(unique(xreact_summary_df$probe_id))
n_genomes_against <- length(unique(hits_against$target_id))

# --- Load coverage data ---
cov_summary_df <- read.delim(params$coverage_summary_file, stringsAsFactors = FALSE)
depth_intervals <- read.delim(params$coverage_depth_file, stringsAsFactors = FALSE)

# Interval expansion helper (from probe_coverage.Rmd)
expand_intervals <- function(interval_df, max_points = 2000) {
  if (nrow(interval_df) == 0) {
    return(data.frame(reference_id = character(), position = integer(), depth = numeric()))
  }
  ref_id <- interval_df$reference_id[1]
  total_len <- max(interval_df$end) - min(interval_df$start) + 1L
  if (total_len <= max_points) {
    positions <- unlist(mapply(seq, interval_df$start, interval_df$end, SIMPLIFY = FALSE))
    depths <- rep(interval_df$depth, interval_df$end - interval_df$start + 1L)
    return(data.frame(reference_id = ref_id, position = positions, depth = depths))
  }
  genome_start <- min(interval_df$start)
  genome_end <- max(interval_df$end)
  bin_size <- (genome_end - genome_start + 1) / max_points
  bin_edges <- genome_start + bin_size * (0:max_points)
  bin_depth_sum <- numeric(max_points)
  bin_width <- numeric(max_points)
  for (i in seq_len(nrow(interval_df))) {
    iv_start <- interval_df$start[i]
    iv_end   <- interval_df$end[i]
    iv_depth <- interval_df$depth[i]
    first_bin <- max(1L, ceiling((iv_start - genome_start) / bin_size))
    last_bin  <- min(max_points, ceiling((iv_end - genome_start) / bin_size))
    if (first_bin > max_points || last_bin < 1L) next
    first_bin <- max(1L, first_bin)
    last_bin  <- min(max_points, last_bin)
    for (b in first_bin:last_bin) {
      overlap_start <- max(iv_start, bin_edges[b])
      overlap_end   <- min(iv_end, bin_edges[b + 1] - 1)
      overlap_len   <- overlap_end - overlap_start + 1
      if (overlap_len > 0) {
        bin_depth_sum[b] <- bin_depth_sum[b] + iv_depth * overlap_len
        bin_width[b]     <- bin_width[b] + overlap_len
      }
    }
  }
  avg_depth <- ifelse(bin_width > 0, bin_depth_sum / bin_width, 0)
  bin_centers <- (bin_edges[1:max_points] + bin_edges[2:(max_points + 1)]) / 2
  data.frame(reference_id = ref_id, position = bin_centers, depth = avg_depth)
}

has_multi <- nchar(params$coverage_multi_mapping_file) > 0 &&
  file.exists(params$coverage_multi_mapping_file)
if (has_multi) {
  multi_df <- read.delim(params$coverage_multi_mapping_file, stringsAsFactors = FALSE)
} else {
  multi_df <- data.frame(probe_id = character(), num_targets = integer(), targets = character())
}

n_targets <- nrow(cov_summary_df)
is_small  <- n_targets <= 20
is_medium <- n_targets > 20 & n_targets <= 100
is_large  <- n_targets > 100

proximity_bp <- params$coverage_proximity

has_params <- nchar(params$params_file) > 0 && file.exists(params$params_file)
if (has_params) {
  params_df <- read.delim(params$params_file, stringsAsFactors = FALSE)
}

# --- Load individual target coverage data (optional) ---
has_individual_cov <- nchar(params$individual_coverage_file) > 0 &&
  file.exists(params$individual_coverage_file)
if (has_individual_cov) {
  indiv_cov_df <- read.delim(params$individual_coverage_file,
                              stringsAsFactors = FALSE)
}

# --- Load refinement summary (optional) ---
has_refine <- nchar(params$refine_summary_file) > 0 &&
  file.exists(params$refine_summary_file)
if (has_refine) {
  refine_df <- read.delim(params$refine_summary_file,
                           stringsAsFactors = FALSE)
  n_refine_iters <- nrow(refine_df) - 1L
  refine_initial_below <- refine_df$num_below_threshold[1]
  refine_final_below   <- refine_df$num_below_threshold[nrow(refine_df)]
}
```

**Generated:** `r Sys.time()`

---

```{r build_section_header, eval=has_build_stats, results='asis'}
cat("\n# Build Pipeline {.tabset}\n")
```

```{r build_section_header2, eval=!has_build_stats, results='asis'}
cat("")
```

```{r build_summary_header, eval=has_build_stats, results='asis'}
cat("\n## Pipeline Summary\n")
```

```{r build_summary_table, eval=has_build_stats, results='asis'}
display <- build_stats %>%
  mutate(
    Step = step,
    Sequences = format(sequences, big.mark = ","),
    Bases = format(bases, big.mark = ","),
    `Dropped (seqs)` = ifelse(is.na(dropped_seqs), "-", format(dropped_seqs, big.mark = ",")),
    `Retained (%)` = ifelse(is.na(pct_retained), "-", paste0(pct_retained, "%"))
  ) %>%
  select(Step, Sequences, Bases, `Dropped (seqs)`, `Retained (%)`)

knitr::kable(display, align = "lrrrr")
```

```{r build_retention_text, eval=has_build_stats, results='asis'}
cat(sprintf(
  "\n**Overall retention:** %s probes from %s input sequences (%s%%).\n",
  format(build_stats$sequences[nrow(build_stats)], big.mark = ","),
  format(build_stats$sequences[1], big.mark = ","),
  round(100 * build_stats$sequences[nrow(build_stats)] / max(build_stats$sequences[1], 1), 1)
))
```

```{r build_seq_header, eval=has_build_stats, results='asis'}
cat("\n## Sequences per Step\n")
```

```{r build_seq_chart, eval=has_build_stats}
ggplot(build_stats, aes(x = step, y = sequences)) +
  geom_col(fill = "#2c7fb8", width = 0.6) +
  geom_text(aes(label = format(sequences, big.mark = ",")),
            vjust = -0.5, size = 4) +
  labs(x = "Pipeline Step", y = "Number of Sequences",
       title = "Sequence Count Through Build Pipeline") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r build_bases_header, eval=has_build_stats, results='asis'}
cat("\n## Total Bases per Step\n")
```

```{r build_bases_chart, eval=has_build_stats}
ggplot(build_stats, aes(x = step, y = bases)) +
  geom_col(fill = "#41ae76", width = 0.6) +
  geom_text(aes(label = format(bases, big.mark = ",")),
            vjust = -0.5, size = 4) +
  labs(x = "Pipeline Step", y = "Total Bases",
       title = "Total Bases Through Build Pipeline") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

---

# Probe Coverage

**Targets analyzed:** `r n_targets` | **Fully tiled (100%):** `r sum(cov_summary_df$pct_covered_1x >= 100)` | **Well tiled (>=90%):** `r sum(cov_summary_df$pct_covered_1x >= 90)` | **Multi-mapping probes:** `r nrow(multi_df)` | **Proximity:** `r proximity_bp` bp

## Summary Table

```{r cov-summary-table}
display_df <- cov_summary_df %>%
  arrange(pct_covered_1x) %>%
  mutate(
    ref_length_fmt = format(ref_length, big.mark = ","),
    covered_bases_fmt = format(covered_bases, big.mark = ","),
    pct_covered_1x_fmt = sprintf("%.1f%%", pct_covered_1x),
    mean_depth_fmt = sprintf("%.1f", mean_depth),
    median_depth_fmt = sprintf("%.1f", median_depth),
    pct_covered_2x_fmt = sprintf("%.1f%%", pct_covered_2x),
    pct_covered_5x_fmt = sprintf("%.1f%%", pct_covered_5x),
    pct_covered_10x_fmt = sprintf("%.1f%%", pct_covered_10x),
    max_gap_length_fmt = format(max_gap_length, big.mark = ","),
    pct_near_probe_fmt = sprintf("%.1f%%", pct_near_probe)
  )

if (is_small) {
  display_df %>%
    select(reference_id, ref_length_fmt, covered_bases_fmt, pct_covered_1x_fmt,
           mean_depth_fmt, median_depth_fmt, pct_covered_2x_fmt, pct_covered_5x_fmt,
           pct_covered_10x_fmt, max_gap_length_fmt, num_gaps, pct_near_probe_fmt) %>%
    knitr::kable(
      col.names = c("Reference", "Length", "Covered", "%\u22651X", "Mean Depth",
                     "Median Depth", "%\u22652X", "%\u22655X", "%\u226510X",
                     "Max Gap", "Gaps", "% Near Probe"),
      align = "lrrrrrrrrrrr"
    )
} else {
  cov_summary_df %>%
    arrange(pct_covered_1x) %>%
    select(reference_id, ref_length, covered_bases, pct_covered_1x,
           mean_depth, median_depth, pct_covered_2x, pct_covered_5x,
           pct_covered_10x, max_gap_length, num_gaps, pct_near_probe) %>%
    datatable(
      colnames = c("Reference", "Length", "Covered", "%\u22651X", "Mean Depth",
                    "Median Depth", "%\u22652X", "%\u22655X", "%\u226510X",
                    "Max Gap", "Gaps", "% Near Probe"),
      options = list(pageLength = 15, scrollX = TRUE, order = list(list(3, "asc"))),
      rownames = FALSE,
      filter = "top"
    ) %>%
      formatRound(columns = c("mean_depth", "median_depth"), digits = 1) %>%
      formatRound(columns = c("pct_covered_1x", "pct_covered_2x", "pct_covered_5x",
                               "pct_covered_10x", "pct_near_probe"), digits = 1) %>%
      formatCurrency(columns = c("ref_length", "covered_bases", "max_gap_length"),
                     currency = "", digits = 0)
}
```

## Coverage Breadth

```{r cov-breadth, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 5, fig.width=if (is_large) 10 else 8}
cov_summary_df <- cov_summary_df %>%
  mutate(
    coverage_class = case_when(
      pct_covered_1x >= 95 ~ "Excellent (>=95%)",
      pct_covered_1x >= 80 ~ "Good (>=80%)",
      TRUE ~ "Poor (<80%)"
    ),
    coverage_class = factor(coverage_class,
                            levels = c("Excellent (>=95%)", "Good (>=80%)", "Poor (<80%)"))
  )

quality_colors <- c("Excellent (>=95%)" = "#27ae60",
                     "Good (>=80%)" = "#f39c12",
                     "Poor (<80%)" = "#e74c3c")

if (is_small) {
  ggplot(cov_summary_df, aes(x = reorder(reference_id, pct_covered_1x),
                          y = pct_covered_1x, fill = coverage_class)) +
    geom_col(width = 0.7) +
    geom_hline(yintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.3) +
    coord_flip() +
    scale_fill_manual(values = quality_colors) +
    labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
         title = "Probe Tiling Breadth per Target") +
    theme(legend.position = "bottom")
} else if (is_medium) {
  ggplot(cov_summary_df, aes(x = reorder(reference_id, pct_covered_1x),
                          y = pct_covered_1x, fill = coverage_class)) +
    geom_col(width = 0.9) +
    geom_hline(yintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.3) +
    coord_flip() +
    scale_fill_manual(values = quality_colors) +
    labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
         title = "Probe Tiling Breadth per Target") +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))
} else {
  ggplot(cov_summary_df, aes(x = pct_covered_1x, fill = coverage_class)) +
    geom_histogram(binwidth = 2, boundary = 0, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = quality_colors) +
    geom_vline(xintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.5) +
    labs(x = "% Bases Covered (>= 1X)", y = "Number of Targets", fill = "Quality",
         title = "Distribution of Probe Tiling Breadth",
         subtitle = sprintf("%d targets | Dashed lines: 80%% and 95%%", n_targets)) +
    theme(legend.position = "bottom")
}
```

```{r cov-worst, eval=is_large}
worst <- cov_summary_df %>%
  arrange(pct_covered_1x) %>%
  head(20) %>%
  select(reference_id, ref_length, pct_covered_1x, mean_depth, max_gap_length, num_gaps)

datatable(worst,
  colnames = c("Reference", "Length", "%\u22651X", "Mean Depth", "Max Gap", "Gaps"),
  options = list(pageLength = 10, dom = "tp"),
  rownames = FALSE
) %>%
  formatRound(columns = c("pct_covered_1x", "mean_depth"), digits = 1) %>%
  formatCurrency(columns = c("ref_length", "max_gap_length"), currency = "", digits = 0)
```

## Individual Target Coverage

```{r indiv-cov-section, eval=has_individual_cov, results='asis'}
cat(paste0(
  "Coverage computed for each target in isolation (minimap2 run once per target ",
  "against that target alone). Probe competition from similar targets is entirely ",
  "eliminated, giving the theoretical maximum coverage achievable for each target ",
  "if it were the only sequence in the panel.\n\n"
))
```

```{r indiv-cov-absent, eval=!has_individual_cov, results='asis'}
cat("*Individual target coverage not computed. Re-run without `--no-individual-targets` to enable.*\n")
```

```{r indiv-cov-breadth, eval=has_individual_cov, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 5, fig.width=if (is_large) 10 else 8}
indiv_n <- nrow(indiv_cov_df)
indiv_small  <- indiv_n <= 20
indiv_medium <- indiv_n > 20 & indiv_n <= 100
indiv_large  <- indiv_n > 100

indiv_cov_df <- indiv_cov_df %>%
  mutate(
    coverage_class = case_when(
      pct_covered_1x >= 95 ~ "Excellent (>=95%)",
      pct_covered_1x >= 80 ~ "Good (>=80%)",
      TRUE ~ "Poor (<80%)"
    ),
    coverage_class = factor(coverage_class,
                            levels = c("Excellent (>=95%)", "Good (>=80%)", "Poor (<80%)"))
  )

quality_colors_indiv <- c("Excellent (>=95%)" = "#27ae60",
                           "Good (>=80%)"      = "#f39c12",
                           "Poor (<80%)"       = "#e74c3c")

if (indiv_small) {
  ggplot(indiv_cov_df, aes(x = reorder(reference_id, pct_covered_1x),
                            y = pct_covered_1x, fill = coverage_class)) +
    geom_col(width = 0.7) +
    geom_hline(yintercept = c(80, 95), linetype = "dashed",
               color = "grey40", linewidth = 0.3) +
    coord_flip() +
    scale_fill_manual(values = quality_colors_indiv) +
    labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
         title = "Individual Probe Tiling Breadth per Target",
         subtitle = "Each target evaluated in isolation (no probe competition)") +
    theme(legend.position = "bottom")
} else if (indiv_medium) {
  ggplot(indiv_cov_df, aes(x = reorder(reference_id, pct_covered_1x),
                            y = pct_covered_1x, fill = coverage_class)) +
    geom_col(width = 0.9) +
    geom_hline(yintercept = c(80, 95), linetype = "dashed",
               color = "grey40", linewidth = 0.3) +
    coord_flip() +
    scale_fill_manual(values = quality_colors_indiv) +
    labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
         title = "Individual Probe Tiling Breadth per Target",
         subtitle = "Each target evaluated in isolation (no probe competition)") +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))
} else {
  ggplot(indiv_cov_df, aes(x = pct_covered_1x, fill = coverage_class)) +
    geom_histogram(binwidth = 2, boundary = 0, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = quality_colors_indiv) +
    geom_vline(xintercept = c(80, 95), linetype = "dashed",
               color = "grey40", linewidth = 0.5) +
    labs(x = "% Bases Covered (>= 1X)", y = "Number of Targets", fill = "Quality",
         title = "Distribution of Individual Probe Tiling Breadth",
         subtitle = sprintf("%d targets | Each evaluated in isolation | Dashed lines: 80%% and 95%%",
                            indiv_n)) +
    theme(legend.position = "bottom")
}
```

```{r indiv-cov-poor-msg, eval=has_individual_cov, results='asis'}
poor_df <- indiv_cov_df %>%
  filter(pct_covered_1x < 80) %>%
  arrange(pct_covered_1x) %>%
  select(Name = reference_id, `Length (bp)` = ref_length,
         `% >= 1X` = pct_covered_1x, `Mean Depth` = mean_depth)

if (nrow(poor_df) > 0) {
  cat(sprintf("\n**%d target%s below 80%% 1X coverage:**\n\n",
              nrow(poor_df), if (nrow(poor_df) == 1) "" else "s"))
} else {
  cat("\n*All targets meet the 80% 1X coverage threshold.*\n")
}
```

```{r indiv-cov-poor-table, eval=has_individual_cov && nrow(poor_df) > 0}
DT::datatable(poor_df, rownames = FALSE, options = list(
  pageLength = 25, scrollX = TRUE, dom = "tip"
)) %>%
  DT::formatRound(columns = c("% >= 1X", "Mean Depth"), digits = 1)
```

```{r indiv-cov-comparison, eval=has_individual_cov && nrow(indiv_cov_df) == n_targets, fig.height=5, fig.width=7}
compare_df <- cov_summary_df %>%
  select(reference_id, pangenome = pct_covered_1x) %>%
  inner_join(indiv_cov_df %>% select(reference_id, individual = pct_covered_1x),
             by = "reference_id") %>%
  mutate(gain = individual - pangenome)

ggplot(compare_df, aes(x = pangenome, y = individual)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "grey60", linewidth = 0.5) +
  geom_point(aes(color = gain), alpha = 0.7, size = 2) +
  scale_color_gradient2(low = "#2c3e50", mid = "#f39c12", high = "#e74c3c",
                        midpoint = 10,
                        name = "Gain (pp)") +
  labs(x = "Pangenome coverage (%\u22651X)",
       y = "Individual coverage (%\u22651X)",
       title = "Probe Competition Effect",
       subtitle = "Points above the diagonal benefit from individual evaluation") +
  coord_fixed(xlim = c(0, 100), ylim = c(0, 100))
```

## Tiered Coverage

```{r cov-tiered, fig.height=if (is_small) max(3, n_targets * 0.5) else if (is_medium) 6 else 5}
tiered <- cov_summary_df %>%
  select(reference_id, pct_covered_1x, pct_covered_2x, pct_covered_5x, pct_covered_10x) %>%
  pivot_longer(cols = starts_with("pct_covered_"),
               names_to = "threshold_tier", values_to = "pct") %>%
  mutate(threshold_tier = case_when(
    threshold_tier == "pct_covered_1x"  ~ ">=1X",
    threshold_tier == "pct_covered_2x"  ~ ">=2X",
    threshold_tier == "pct_covered_5x"  ~ ">=5X",
    threshold_tier == "pct_covered_10x" ~ ">=10X"
  ),
  threshold_tier = factor(threshold_tier, levels = c(">=1X", ">=2X", ">=5X", ">=10X")))

tier_colors <- c(">=1X" = "#3498db", ">=2X" = "#2ecc71",
                  ">=5X" = "#f39c12", ">=10X" = "#e74c3c")

if (is_small) {
  ggplot(tiered, aes(x = reorder(reference_id, pct), y = pct, fill = threshold_tier)) +
    geom_col(position = "dodge", width = 0.7) +
    coord_flip() +
    scale_fill_manual(values = tier_colors) +
    labs(x = NULL, y = "% Bases Covered", fill = "Depth Threshold",
         title = "Tiered Probe Coverage per Target") +
    theme(legend.position = "bottom")
} else if (is_medium) {
  ggplot(tiered, aes(x = reorder(reference_id, pct), y = pct, fill = threshold_tier)) +
    geom_col(position = "dodge", width = 0.9) +
    coord_flip() +
    scale_fill_manual(values = tier_colors) +
    labs(x = NULL, y = "% Bases Covered", fill = "Depth Threshold",
         title = "Tiered Probe Coverage per Target") +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))
} else {
  ggplot(tiered, aes(x = threshold_tier, y = pct, fill = threshold_tier)) +
    geom_boxplot(alpha = 0.7, outlier.size = 0.5) +
    geom_jitter(width = 0.2, alpha = 0.1, size = 0.3) +
    scale_fill_manual(values = tier_colors) +
    labs(x = "Depth Threshold", y = "% Bases Covered",
         title = "Distribution of Tiered Probe Coverage",
         subtitle = sprintf("%d targets", n_targets)) +
    theme(legend.position = "none")
}
```

## Gap Analysis

```{r cov-gap, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 4}
gap_data <- cov_summary_df %>% filter(num_gaps > 0)

if (nrow(gap_data) == 0) {
  cat("All targets are fully tiled with no gaps.")
} else if (is_small) {
  ggplot(gap_data, aes(x = reorder(reference_id, max_gap_length), y = max_gap_length)) +
    geom_segment(aes(xend = reference_id, y = 0, yend = max_gap_length),
                 color = "#e74c3c", linewidth = 1) +
    geom_point(size = 4, color = "#e74c3c") +
    geom_text(aes(label = sprintf("%s bp (%d gaps)",
                                   format(max_gap_length, big.mark = ","), num_gaps)),
              hjust = -0.1, size = 3.5) +
    coord_flip() +
    labs(x = NULL, y = "Largest Gap (bp)",
         title = "Largest Uncovered Stretch per Target",
         subtitle = "Targets with no gaps are omitted") +
    expand_limits(y = max(gap_data$max_gap_length) * 1.3)
} else if (is_medium) {
  ggplot(gap_data, aes(x = reorder(reference_id, max_gap_length), y = max_gap_length)) +
    geom_col(fill = "#e74c3c", width = 0.9, alpha = 0.8) +
    coord_flip() +
    labs(x = NULL, y = "Largest Gap (bp)",
         title = "Largest Uncovered Stretch per Target",
         subtitle = sprintf("%d of %d targets have gaps", nrow(gap_data), n_targets)) +
    theme(axis.text.y = element_text(size = 5))
} else {
  gap_data <- gap_data %>% arrange(desc(max_gap_length)) %>%
    mutate(rank = row_number())
  ggplot(gap_data, aes(x = max_gap_length, y = reorder(factor(rank), -rank))) +
    geom_col(fill = "#e74c3c", alpha = 0.7, width = 1) +
    labs(x = "Largest Gap (bp)", y = NULL,
         title = "Gap Sizes Across All Targets",
         subtitle = sprintf("%d of %d targets have gaps (sorted by gap size, largest at top)",
                            nrow(gap_data), n_targets)) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major.y = element_blank())
}
```

## Gap Details

```{r gap-details-load, include=FALSE}
has_gap_details <- nchar(params$gap_details_file) > 0 &&
  file.exists(params$gap_details_file)
gap_min_bp <- as.integer(params$gap_min_length)
if (has_gap_details) {
  gap_df <- read.delim(params$gap_details_file, stringsAsFactors = FALSE,
                       colClasses = c(gap_sequence = "character"))
  has_indiv_gap_cols <- "gap_type" %in% names(gap_df) &&
    any(nchar(gap_df$gap_type) > 0)
}
```

```{r gap-details-summary, results='asis'}
if (!has_gap_details || nrow(gap_df) == 0) {
  cat(sprintf("*No gaps ≥ %d bp found.*\n", gap_min_bp))
} else {
  n_gaps <- nrow(gap_df)
  n_targets_with_gaps <- length(unique(gap_df$target_id))
  cat(sprintf(
    "**%d gaps ≥ %d bp** across **%d targets**.\n\n",
    n_gaps, gap_min_bp, n_targets_with_gaps
  ))
  if (has_indiv_gap_cols) {
    n_true <- sum(gap_df$gap_type == "true_gap", na.rm = TRUE)
    n_mm   <- sum(gap_df$gap_type == "multimapper_gap", na.rm = TRUE)
    cat(sprintf(
      "- **%d true gaps** — no probe aligns to this region even in isolation\n",
      n_true
    ))
    cat(sprintf(
      "- **%d multi-mapper gaps** — a probe exists but was outcompeted in combined mapping\n\n",
      n_mm
    ))
  }
  gap_filename <- basename(params$gap_details_file)
  cat(sprintf(
    "*Full gap details (including sequences): [%s](%s)*\n\n",
    gap_filename, gap_filename
  ))
}
```

```{r gap-details-table, eval=has_gap_details && nrow(gap_df) > 0}
display_cols <- c("target_id", "gap_start", "gap_end", "gap_length",
                  "is_terminal", "gc_content", "dust_score")
if (has_indiv_gap_cols) {
  display_cols <- c(display_cols, "gap_type", "individual_coverage")
}

top_gaps <- gap_df[seq_len(min(20, nrow(gap_df))), display_cols, drop = FALSE]
top_gaps$gc_content <- round(top_gaps$gc_content, 3)
top_gaps$dust_score <- round(top_gaps$dust_score, 3)
if ("individual_coverage" %in% names(top_gaps)) {
  top_gaps$individual_coverage <- round(top_gaps$individual_coverage, 3)
}

dt <- DT::datatable(
  top_gaps,
  caption = sprintf(
    "Top %d gaps by length (full file includes gap sequences)",
    nrow(top_gaps)
  ),
  rownames = FALSE,
  options = list(scrollX = TRUE, pageLength = 20, dom = "t"),
  class = "compact"
)

if (has_indiv_gap_cols && "gap_type" %in% names(top_gaps)) {
  type_col_idx <- which(names(top_gaps) == "gap_type") - 1L
  dt <- DT::formatStyle(
    dt, "gap_type",
    backgroundColor = DT::styleEqual(
      c("true_gap", "multimapper_gap"),
      c("#f9cece", "#fde9cc")
    )
  )
}
dt
```

```{r gap-combined-vs-individual, eval=has_gap_details && has_individual_cov && has_indiv_gap_cols && nrow(gap_df) > 0, fig.height=4, fig.width=7}
scatter_df <- cov_summary_df %>%
  inner_join(
    indiv_cov_df %>% select(reference_id, individual = pct_covered_1x),
    by = "reference_id"
  ) %>%
  mutate(
    has_large_gap = reference_id %in% gap_df$target_id,
    label = if_else(has_large_gap, reference_id, NA_character_)
  )

lim <- c(0, 100)
ggplot(scatter_df, aes(x = pct_covered_1x, y = individual, color = has_large_gap)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_point(alpha = 0.7, size = 2.5) +
  ggrepel::geom_text_repel(aes(label = label), size = 3, max.overlaps = 15,
                            na.rm = TRUE) +
  scale_color_manual(
    values = c("FALSE" = "#3498db", "TRUE" = "#e74c3c"),
    labels = c("No large gap", "Has large gap"),
    name = NULL
  ) +
  coord_fixed(xlim = lim, ylim = lim) +
  labs(
    x = "Combined coverage (%) — real-assay behavior",
    y = "Individual coverage (%) — best possible",
    title = "Combined vs. Individual Target Coverage",
    subtitle = "Points above the diagonal: probes exist but are outcompeted in combined mapping"
  )
```

## Pangenome Probe Depth

Targets concatenated end-to-end, sorted by name.`r if (n_targets <= 12) " Vertical grey lines mark target boundaries." else ""`

```{r cov-pangenome, fig.height=4, fig.width=10}
if (nrow(depth_intervals) > 0) {
  target_order <- cov_summary_df %>% arrange(reference_id)
  offsets <- target_order %>%
    mutate(offset = cumsum(lag(ref_length, default = 0)))
  pan_intervals <- depth_intervals %>%
    inner_join(offsets %>% select(reference_id, offset), by = "reference_id") %>%
    mutate(pan_start = start + offset, pan_end = end + offset) %>%
    arrange(pan_start)
  total_len <- sum(target_order$ref_length)
  n_points <- min(2000, total_len)
  sample_pos <- seq(1, total_len, length.out = n_points)
  idx <- findInterval(sample_pos, pan_intervals$pan_start)
  pan_depth <- ifelse(idx > 0 & sample_pos <= pan_intervals$pan_end[pmax(idx, 1)],
                      pan_intervals$depth[pmax(idx, 1)], 0)
  pan_df <- data.frame(pan_position = sample_pos, depth = pan_depth)
  boundaries <- offsets %>%
    mutate(boundary = offset + ref_length) %>%
    filter(row_number() < n())

  pct_pangenome_1x <- 100 * sum(cov_summary_df$covered_bases) /
    sum(cov_summary_df$ref_length)

  ggplot(pan_df, aes(x = pan_position, y = depth)) +
    geom_area(fill = "#3498db", alpha = 0.4) +
    geom_line(color = "#2c3e50", linewidth = 0.2) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.3) +
    geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.3) +
    { if (nrow(boundaries) > 0 && n_targets <= 12)
        geom_vline(data = boundaries, aes(xintercept = boundary),
                   color = "grey50", linetype = "solid", linewidth = 0.3) } +
    labs(x = "Pangenome Position", y = "Probe Depth",
         title = "Probe Depth Across Concatenated Targets",
         subtitle = sprintf("%d targets | %.1f%% pangenome \u22651X | Dashed lines: 1X (green), 5X (orange)",
                            nrow(target_order), pct_pangenome_1x)) +
    theme(axis.text.x = element_text(size = 7))
}
```

## Probe Depth Profiles {.tabset}

```{r cov-depth-overview-calc, include=FALSE}
n_facet <- if (is_small) n_targets else min(12, n_targets)
n_detail <- if (is_small) n_targets else min(10, n_targets)
show_detail <- n_targets <= 100
```

```{r cov-depth-overview, fig.height=max(3, ceiling(min(n_facet, 12) / 3) * 2.5), fig.width=10}
if (nrow(depth_intervals) > 0) {
  worst_refs <- cov_summary_df %>%
    arrange(pct_covered_1x) %>%
    head(n_facet) %>%
    pull(reference_id)

  if (n_targets > n_facet) {
    overview_subtitle <- sprintf("Showing %d worst-covered of %d targets | Dashed lines: 1X (green), 5X (orange)",
                                  n_facet, n_targets)
  } else {
    overview_subtitle <- "Dashed lines: 1X (green), 5X (orange)"
  }

  plot_data <- do.call(rbind, lapply(worst_refs, function(rid) {
    expanded <- expand_intervals(depth_intervals %>% filter(reference_id == rid), max_points = 500)
    if (nrow(expanded) == 0) return(NULL)
    expanded
  }))

  ggplot(plot_data, aes(x = position, y = depth)) +
    geom_area(fill = "#3498db", alpha = 0.4) +
    geom_line(color = "#2c3e50", linewidth = 0.3) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.3) +
    geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.3) +
    facet_wrap(~reference_id, scales = "free", ncol = 3) +
    labs(x = "Position", y = "Probe Depth",
         title = "Probe Depth Across Targets",
         subtitle = overview_subtitle) +
    theme(strip.text = element_text(size = 8),
          axis.text = element_text(size = 6))
}
```

```{r cov-depth-detail, results='asis', fig.width=8, fig.height=3, eval=show_detail}
if (nrow(depth_intervals) > 0 && nrow(cov_summary_df) > 0) {
  ordered <- cov_summary_df %>% arrange(pct_covered_1x)
  if (n_targets > n_detail) {
    cat(sprintf('\n\n*Showing detail for the %d worst-covered targets (of %d total).*\n\n',
                n_detail, n_targets))
  }
  for (i in seq_len(n_detail)) {
    ref_id <- ordered$reference_id[i]
    ref_intervals_i <- depth_intervals %>% filter(reference_id == ref_id)
    ref_data <- expand_intervals(ref_intervals_i)
    if (nrow(ref_data) == 0) next
    s <- ordered[i, ]
    cat(sprintf('\n\n<details><summary><strong>%s</strong> &mdash; %.1f%% covered, %.1fX mean, %s bp, %d gaps (max %s bp)</summary>\n\n',
                ref_id, s$pct_covered_1x, s$mean_depth,
                format(s$ref_length, big.mark = ","),
                s$num_gaps, format(s$max_gap_length, big.mark = ",")))
    zero_intervals <- ref_intervals_i %>% filter(depth == 0)
    p <- ggplot(ref_data, aes(x = position, y = depth)) +
      geom_area(fill = "#3498db", alpha = 0.3) +
      geom_line(color = "#2c3e50", linewidth = 0.4) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.5) +
      geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.5)
    if (nrow(zero_intervals) > 0) {
      p <- p + geom_segment(data = zero_intervals,
                            aes(x = start, xend = end, y = 0, yend = 0),
                            color = "darkgreen", linewidth = 2, alpha = 0.5,
                            inherit.aes = FALSE)
    }
    p <- p +
      labs(x = "Genome Position", y = "Probe Depth",
           title = sprintf("Probe Depth: %s", ref_id),
           subtitle = sprintf("%.1f%% covered | %.1fX mean | %.1f median | %s bp",
                              s$pct_covered_1x, s$mean_depth, s$median_depth,
                              format(s$ref_length, big.mark = ",")),
           caption = if (nrow(zero_intervals) > 0) "Green bars: uncovered regions" else NULL) +
      theme_minimal(base_size = 12)
    print(p)
    cat('\n\n</details>\n\n')
  }
}
```

```{r cov-depth-large-notice, results='asis', eval=!show_detail}
cat(sprintf('\n\n*Individual depth plots are omitted for large panels (%d targets). See the faceted overview and pangenome plot above.*\n\n', n_targets))
```

## Proximity Coverage

Pull-down zone: `r proximity_bp` bp around each probe.

```{r cov-proximity, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 5}
prox_subtitle_small <- sprintf("Blue: directly covered | Purple: within %d bp pull-down zone", proximity_bp)

if (is_small) {
  ggplot(cov_summary_df, aes(x = reorder(reference_id, pct_near_probe), y = pct_near_probe)) +
    geom_col(fill = "#8e44ad", width = 0.7, alpha = 0.8) +
    geom_col(aes(y = pct_covered_1x), fill = "#3498db", width = 0.7, alpha = 0.5) +
    coord_flip() +
    labs(x = NULL, y = "% Target Bases",
         title = "Direct Coverage vs Proximity Coverage",
         subtitle = prox_subtitle_small) +
    theme(legend.position = "bottom")
} else if (is_medium) {
  ggplot(cov_summary_df, aes(x = reorder(reference_id, pct_near_probe), y = pct_near_probe)) +
    geom_col(fill = "#8e44ad", width = 0.9, alpha = 0.8) +
    geom_col(aes(y = pct_covered_1x), fill = "#3498db", width = 0.9, alpha = 0.5) +
    coord_flip() +
    labs(x = NULL, y = "% Target Bases",
         title = "Direct Coverage vs Proximity Coverage",
         subtitle = prox_subtitle_small) +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))
} else {
  ggplot(cov_summary_df, aes(x = pct_covered_1x, y = pct_near_probe)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(alpha = 0.5, color = "#8e44ad", size = 1.5) +
    labs(x = "% Directly Covered (>=1X)",
         y = sprintf("%% Within %d bp Pull-Down Zone", proximity_bp),
         title = "Direct Coverage vs Proximity Coverage",
         subtitle = sprintf("%d targets | Diagonal: perfect agreement", n_targets)) +
    coord_fixed(xlim = c(0, 100), ylim = c(0, 100))
}
```

## Multi-Mapping Probes

`r nrow(multi_df)` probes map to more than one target sequence.

```{r cov-multi-mapping, eval=(nrow(multi_df) > 0)}
if (nrow(multi_df) <= 20) {
  multi_df %>%
    arrange(desc(num_targets)) %>%
    knitr::kable(
      col.names = c("Probe ID", "# Targets", "Targets"),
      align = "lrl"
    )
} else {
  multi_df %>%
    arrange(desc(num_targets)) %>%
    datatable(
      colnames = c("Probe ID", "# Targets", "Targets"),
      options = list(pageLength = 15, order = list(list(1, "desc"))),
      rownames = FALSE,
      filter = "top"
    )
}
```

---

```{r refine-section-header, eval=has_refine, results='asis'}
cat("\n# Refinement Summary\n")
```

```{r refine-stats, eval=has_refine, results='asis'}
if (n_refine_iters == 0) {
  cat(sprintf(
    paste0("Refinement was requested but no iterations were needed: all %d",
           " targets already met the coverage threshold.\n"),
    refine_df$num_targets[1]
  ))
} else {
  cat(sprintf(
    "**%d refinement iteration%s** ran. ",
    n_refine_iters,
    if (n_refine_iters == 1) "" else "s"
  ))
  cat(sprintf(
    "Initial targets below threshold: **%d**. ",
    refine_initial_below
  ))
  if (refine_final_below == 0) {
    cat("All targets are now covered.\n")
  } else {
    cat(sprintf("Targets remaining below threshold: **%d**.\n",
                refine_final_below))
  }
}
```

```{r refine-table, eval=has_refine}
display_refine <- refine_df %>%
  mutate(
    step = ifelse(step == "initial", "Initial", paste("Iteration", step)),
    pct_covered = sprintf("%.1f%%", pct_covered)
  )

knitr::kable(
  display_refine,
  col.names = c("Step", "Targets Analyzed", "Below Threshold", "% Covered"),
  align = "lrrr"
)
```

```{r refine-chart, eval=has_refine && n_refine_iters > 0, fig.height=4, fig.width=7}
plot_refine <- refine_df %>%
  mutate(
    step_label = ifelse(step == "initial", "Initial",
                        paste("Iter", step)),
    step_label = factor(step_label, levels = step_label),
    covered = num_targets - num_below_threshold
  ) %>%
  select(step_label, covered, num_below_threshold) %>%
  pivot_longer(cols = c(covered, num_below_threshold),
               names_to = "category", values_to = "count") %>%
  mutate(
    category = ifelse(category == "covered",
                      "Covered", "Below threshold"),
    category = factor(category,
                      levels = c("Below threshold", "Covered"))
  )

ggplot(plot_refine, aes(x = step_label, y = count, fill = category)) +
  geom_col(width = 0.6) +
  scale_fill_manual(values = c("Covered" = "#27ae60",
                                "Below threshold" = "#e74c3c")) +
  geom_text(data = plot_refine %>% filter(category == "Below threshold"),
            aes(label = ifelse(count > 0, count, "")),
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold", size = 4) +
  labs(x = NULL, y = "Number of Targets", fill = NULL,
       title = "Coverage Progress Across Refinement Steps",
       subtitle = "Red: targets below threshold | Green: targets at or above threshold") +
  theme(legend.position = "bottom")
```

---

# Cross-Reactivity

```{r xreact-summary-table, results='asis'}
info <- data.frame(
  Parameter = c("Homology threshold", "Total probes"),
  Value = c(paste0(threshold, "%"), n_probes_xreact),
  stringsAsFactors = FALSE
)

if (has_self) {
  n_self_flagged <- sum(summary_self$num_hits_above_threshold > 0)
  info <- rbind(info, data.frame(
    Parameter = c("Probes with self-homology hits", "Self-mode hits"),
    Value = c(paste0(n_self_flagged, " / ", n_probes_xreact,
                     " (", round(100 * n_self_flagged / max(n_probes_xreact, 1), 1), "%)"),
              nrow(hits_self))
  ))
}

if (has_against) {
  n_against_flagged <- sum(summary_against$num_hits_above_threshold > 0)
  info <- rbind(info, data.frame(
    Parameter = c("Probes with cross-reactive genome hits",
                  "Genomes with hits", "Against-mode hits"),
    Value = c(paste0(n_against_flagged, " / ", n_probes_xreact,
                     " (", round(100 * n_against_flagged / max(n_probes_xreact, 1), 1), "%)"),
              n_genomes_against,
              nrow(hits_against))
  ))
}

knitr::kable(info, col.names = c("Parameter", "Value"), align = "lr")
```

```{r self-section-header, eval=has_self, results='asis'}
cat("\n## Self-Homology (Probe vs Probe)\n")
```

```{r self-heatmap, eval=has_self && n_probes_xreact < 1000, out.width="100%"}
all_probe_ids <- sort(unique(xreact_summary_df$probe_id))
n <- length(all_probe_ids)

mat <- matrix(NA_real_, nrow = n, ncol = n,
              dimnames = list(all_probe_ids, all_probe_ids))
hover_mat <- matrix("", nrow = n, ncol = n,
                    dimnames = list(all_probe_ids, all_probe_ids))

if (nrow(hits_self) > 0) {
  for (i in seq_len(nrow(hits_self))) {
    row <- hits_self[i, ]
    ri <- match(row$target_id, all_probe_ids)
    ci <- match(row$probe_id, all_probe_ids)
    if (!is.na(ri) && !is.na(ci)) {
      if (is.na(mat[ri, ci]) || row$homology_pct > mat[ri, ci]) {
        mat[ri, ci] <- row$homology_pct
        hover_mat[ri, ci] <- paste0(
          "Probe: ", row$probe_id, "<br>",
          "Target: ", row$target_id, "<br>",
          "Homology: ", round(row$homology_pct, 1), "%<br>",
          "Identity: ", round(row$identity_pct, 1), "%<br>",
          "Query coverage: ", round(row$query_coverage_pct, 1), "%"
        )
      }
    }
  }
}

show_labels <- n <= 20

plot_ly(
  z = mat,
  x = colnames(mat),
  y = rownames(mat),
  type = "heatmap",
  colorscale = list(c(0, "steelblue"), c(0.5, "orange"), c(1, "firebrick")),
  zmin = threshold,
  zmax = 100,
  hoverinfo = "text",
  text = hover_mat,
  showscale = TRUE,
  colorbar = list(title = "Homology %"),
  height = 800,
  width = 1000
) %>%
  layout(
    title = "Self-Homology Matrix",
    xaxis = list(
      title = "Probe",
      showticklabels = show_labels,
      tickangle = if (show_labels) -45 else 0,
      showgrid = FALSE
    ),
    yaxis = list(
      title = "Target probe",
      showticklabels = show_labels,
      autorange = "reversed",
      showgrid = FALSE
    ),
    plot_bgcolor = "white",
    height = NULL,
    xgap = 0.5, ygap = 0.5
  )
```

```{r self-heatmap-skip, eval=has_self && n_probes_xreact >= 1000, results='asis'}
cat(paste0("\n*Heatmap omitted: ", n_probes_xreact,
           " probes exceeds the 1,000 probe limit for interactive display.*\n"))
```

```{r self-density-header, eval=has_self && nrow(hits_self) > 0, results='asis'}
cat("\n### Homology Distribution\n")
```

```{r self-density, eval=has_self && nrow(hits_self) > 0, fig.height=4}
ggplot(summary_self, aes(x = max_homology_pct)) +
  geom_density(fill = "steelblue", alpha = 0.5, color = "steelblue", adjust = 1/4) +
  geom_vline(xintercept = threshold, linetype = "dashed", color = "firebrick", linewidth = 0.8) +
  annotate("text", x = threshold, y = Inf, label = paste0("Threshold: ", threshold, "%"),
           hjust = -0.1, vjust = 1.5, color = "firebrick", size = 3.5) +
  labs(x = "Homology (%)", y = "Density", title = "Maximum Homology for All Probes") +
  theme_minimal(base_size = 14)

ggplot(hits_self, aes(x = homology_pct)) +
  geom_density(fill = "steelblue", alpha = 0.5, color = "steelblue", adjust = 1/2) +
  geom_vline(xintercept = threshold, linetype = "dashed", color = "firebrick", linewidth = 0.8) +
  annotate("text", x = threshold, y = Inf, label = paste0("Threshold: ", threshold, "%"),
           hjust = -0.1, vjust = 1.5, color = "firebrick", size = 3.5) +
  scale_x_continuous(limits = c(max(threshold - 10, 0), 100)) +
  labs(x = "Homology (%)", y = "Density", title = "Self-Homology Distribution for Hits Above Threshold") +
  theme_minimal(base_size = 14)
```

```{r self-table-header, eval=has_self, results='asis'}
cat("\n### All Self-Homology Hits\n")
```

```{r self-table, eval=has_self}
display_self <- hits_self %>%
  select(probe_id, target_id, homology_pct, identity_pct,
         query_coverage_pct, matching_bases, probe_length) %>%
  arrange(desc(homology_pct))

datatable(
  display_self,
  colnames = c("Probe", "Target Probe", "Homology %", "Identity %",
               "Query Cov %", "Matching Bases", "Probe Length"),
  options = list(pageLength = 25, scrollX = TRUE,
                 order = list(list(2, "desc"))),
  rownames = FALSE,
  filter = if (nrow(display_self) > 50) "top" else "none"
) %>%
  formatRound(c("homology_pct", "identity_pct", "query_coverage_pct"), 1)
```

```{r against-section-header, eval=has_against, results='asis'}
cat("\n## Cross-Reactivity (Probe vs Genome)\n")
```

```{r against-no-hits, eval=has_against && nrow(hits_against) == 0, results='asis'}
cat("\n<p style='font-size:1.6em; font-weight:bold; text-align:center; padding:2em 0; color:#2c7bb6;'>No Significant Cross Reactivity Found</p>\n")
```

```{r against-heatmap, eval=has_against && nrow(hits_against) > 0 && n_probes_xreact < 1000, fig.height=8}
probes_with_hits <- sort(unique(hits_against$probe_id))
genomes <- sort(unique(hits_against$target_id))
n_probes_hit <- length(probes_with_hits)
n_genomes_hit <- length(genomes)

if (n_probes_hit > 0 && n_genomes_hit > 0) {
  mat_ag <- matrix(NA_real_, nrow = n_genomes_hit, ncol = n_probes_hit,
                   dimnames = list(genomes, probes_with_hits))
  hover_ag <- matrix("", nrow = n_genomes_hit, ncol = n_probes_hit,
                     dimnames = list(genomes, probes_with_hits))

  for (i in seq_len(nrow(hits_against))) {
    row <- hits_against[i, ]
    ri <- match(row$target_id, genomes)
    ci <- match(row$probe_id, probes_with_hits)
    if (!is.na(ri) && !is.na(ci)) {
      if (is.na(mat_ag[ri, ci]) || row$homology_pct > mat_ag[ri, ci]) {
        mat_ag[ri, ci] <- row$homology_pct
        hover_ag[ri, ci] <- paste0(
          "Probe: ", row$probe_id, "<br>",
          "Genome: ", row$target_id, "<br>",
          "Homology: ", round(row$homology_pct, 1), "%<br>",
          "Identity: ", round(row$identity_pct, 1), "%<br>",
          "Query coverage: ", round(row$query_coverage_pct, 1), "%"
        )
      }
    }
  }

  show_x <- n_probes_hit <= 20
  show_y <- n_genomes_hit <= 20
  plot_height <- max(400, n_genomes_hit * 20 + 150)

  plot_ly(
    z = mat_ag,
    x = colnames(mat_ag),
    y = rownames(mat_ag),
    type = "heatmap",
    colorscale = list(c(0, "steelblue"), c(0.5, "orange"), c(1, "firebrick")),
    zmin = threshold,
    zmax = 100,
    hoverinfo = "text",
    text = hover_ag,
    showscale = TRUE,
    colorbar = list(title = "Homology %"),
    height = plot_height
  ) %>%
    layout(
      title = paste0("Cross-Reactivity Matrix (",
                     n_probes_hit, " probes x ", n_genomes_hit, " genomes)"),
      xaxis = list(
        title = "Probe",
        showticklabels = show_x,
        tickangle = if (show_x) -45 else 0,
        showgrid = FALSE
      ),
      yaxis = list(
        title = "Genome",
        showticklabels = show_y,
        autorange = "reversed",
        showgrid = FALSE
      ),
      plot_bgcolor = "white"
    )
}
```

```{r against-heatmap-skip, eval=has_against && nrow(hits_against) > 0 && n_probes_xreact >= 1000, results='asis'}
cat(paste0("\n*Heatmap omitted: ", n_probes_xreact,
           " probes exceeds the 1,000 probe limit for interactive display.*\n"))
```

```{r against-genome-bar-header, eval=has_against && nrow(hits_against) > 0, results='asis'}
cat("\n### Probes per Genome\n")
```

```{r against-genome-bar, eval=has_against && nrow(hits_against) > 0 && n_genomes_against <= 100, fig.height=max(4, n_genomes_against * 0.3 + 1)}
genome_counts <- hits_against %>%
  group_by(target_id) %>%
  summarise(n_probes = n_distinct(probe_id), .groups = "drop") %>%
  arrange(n_probes) %>%
  mutate(target_id = factor(target_id, levels = target_id))

ggplot(genome_counts, aes(x = n_probes, y = target_id)) +
  geom_col(fill = "steelblue", alpha = 0.8) +
  labs(x = "Number of cross-reactive probes", y = NULL,
       title = "Cross-Reactive Probes per Genome") +
  theme_minimal(base_size = if (n_genomes_against <= 20) 14 else 8)
```

```{r against-genome-bar-large, eval=has_against && nrow(hits_against) > 0 && n_genomes_against > 100, fig.height=5}
genome_counts <- hits_against %>%
  group_by(target_id) %>%
  summarise(n_probes = n_distinct(probe_id), .groups = "drop")

ggplot(genome_counts, aes(x = n_probes)) +
  geom_histogram(fill = "steelblue", alpha = 0.8, color = "white",
                 binwidth = max(1, ceiling(max(genome_counts$n_probes) / 30))) +
  labs(x = "Number of cross-reactive probes per genome", y = "Number of genomes",
       title = paste0("Distribution of Cross-Reactive Probes (",
                      n_genomes_against, " genomes)")) +
  theme_minimal(base_size = 14)
```

```{r against-density-header, eval=has_against && nrow(hits_against) > 0, results='asis'}
cat("\n### Homology Distribution\n")
```

```{r against-density, eval=has_against && nrow(hits_against) > 0, fig.height=4}
ggplot(hits_against, aes(x = homology_pct)) +
  geom_density(fill = "steelblue", alpha = 0.5, color = "steelblue") +
  geom_vline(xintercept = threshold, linetype = "dashed", color = "firebrick", linewidth = 0.8) +
  annotate("text", x = threshold, y = Inf, label = paste0("Threshold: ", threshold, "%"),
           hjust = -0.1, vjust = 1.5, color = "firebrick", size = 3.5) +
  scale_x_continuous(limits = c(max(threshold - 10, 0), 100)) +
  labs(x = "Homology (%)", y = "Density", title = "Cross-Reactivity Homology Distribution") +
  theme_minimal(base_size = 14)
```

```{r against-table-header, eval=has_against && nrow(hits_against) > 0, results='asis'}
cat("\n### All Cross-Reactivity Hits\n")
```

```{r against-table, eval=has_against && nrow(hits_against) > 0}
display_against <- hits_against %>%
  select(probe_id, target_id, homology_pct, identity_pct,
         query_coverage_pct, matching_bases, probe_length) %>%
  arrange(desc(homology_pct))

datatable(
  display_against,
  colnames = c("Probe", "Genome", "Homology %", "Identity %",
               "Query Cov %", "Matching Bases", "Probe Length"),
  options = list(pageLength = 25, scrollX = TRUE,
                 order = list(list(2, "desc"))),
  rownames = FALSE,
  filter = if (nrow(display_against) > 50) "top" else "none"
) %>%
  formatRound(c("homology_pct", "identity_pct", "query_coverage_pct"), 1)
```

---

```{r params-fold, eval=has_params, results='asis'}
cat('<details><summary>Parameters</summary>\n\n')
```

```{r params-table, eval=has_params}
display_params <- params_df %>% select(parameter, value)
knitr::kable(display_params, col.names = c("Parameter", "Value"), align = "ll")
```

```{r params-command, eval=has_params, results='asis'}
cmd_parts <- c("baitbench assess-probes")
for (i in seq_len(nrow(params_df))) {
  flag <- params_df$flag[i]
  value <- params_df$value[i]
  if (value == "none" || value == "false") next
  if (value == "true") {
    cmd_parts <- c(cmd_parts, sprintf("  %s", flag))
  } else {
    cmd_parts <- c(cmd_parts, sprintf("  %s %s", flag, value))
  }
}
cat("```bash\n")
cat(paste(cmd_parts, collapse = " \\\n"), "\n")
cat("```\n")
```

```{r params-fold-end, eval=has_params, results='asis'}
cat('</details>\n\n')
```

```{r build-params-fold, eval=has_build_params, results='asis'}
cat('<details><summary>Build Probes Parameters</summary>\n\n')
```

```{r build-params-table, eval=has_build_params}
display_bp <- build_params_df %>% select(parameter, value)
knitr::kable(display_bp, col.names = c("Parameter", "Value"), align = "ll")
```

```{r build-params-command, eval=has_build_params, results='asis'}
cmd_parts <- c("baitbench build-probes")
for (i in seq_len(nrow(build_params_df))) {
  flag <- build_params_df$flag[i]
  val <- build_params_df$value[i]
  if (!is.na(val) && val != "" && val != "none" && val != "false") {
    if (val == "true") {
      cmd_parts <- c(cmd_parts, flag)
    } else {
      cmd_parts <- c(cmd_parts, paste(flag, val))
    }
  }
}
cat("\n**Command:**\n```\n", paste(cmd_parts, collapse = " \\\n  "), "\n```\n")
```

```{r build-params-fold-end, eval=has_build_params, results='asis'}
cat('</details>\n\n')
```

---

*Generated by [BaitBench](https://github.com/CrossLab/BaitBench) assess-probes*
