---
title: "BaitBench Coverage Curve Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  sweep_file: ""
  sample_ids: !r c()
  swept_params: !r c()
  params_file: ""
---

<style>
.main-container { max-width: 100%; margin-left: 1em; margin-right: 1em; }
table { width: 100%; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 6, out.width = "100%")

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  library(DT)
})

theme_set(theme_minimal(base_size = 14))

# Read data
sweep_df <- read.delim(params$sweep_file, stringsAsFactors = FALSE)

sample_ids <- params$sample_ids
swept_params <- params$swept_params
n_targets <- length(unique(sweep_df$reference_id))

# Read run parameters if available
has_params <- nchar(params$params_file) > 0 && file.exists(params$params_file)
if (has_params) {
  params_df <- read.delim(params$params_file, stringsAsFactors = FALSE)
}

# Detect swept params from data if not explicitly provided
if (length(swept_params) == 0 || (length(swept_params) == 1 && swept_params[1] == "")) {
  swept_params <- character(0)
  if (length(unique(sweep_df$ct)) > 1) swept_params <- c(swept_params, "ct")
  if ("hybridization_temperature" %in% names(sweep_df) && length(unique(sweep_df$hybridization_temperature)) > 1)
    swept_params <- c(swept_params, "hybridization_temperature")
  if (length(unique(sweep_df$capture_fraction)) > 1) swept_params <- c(swept_params, "capture_fraction")
  if (length(unique(sweep_df$num_sequences)) > 1) swept_params <- c(swept_params, "num_sequences")
}

# Format numbers with commas for display (e.g. 10000 -> "10,000")
format_number <- function(x) {
  sapply(x, function(v) {
    if (is.na(v)) return(NA_character_)
    if (v == floor(v)) {
      formatC(as.integer(v), format = "d", big.mark = ",")
    } else {
      as.character(v)
    }
  })
}

# Full display names for swept parameter types
param_full_names <- c(
  ct = "CT",
  hybridization_temperature = "Hybridization Temp (°C)",
  capture_fraction = "Capture Fraction",
  num_sequences = "Num Sequences"
)

# Build combo label from swept params
# When a single param is swept, omits the prefix (e.g. "10,000" instead of "NS=10,000")
build_combo_label <- function(df, swept) {
  if (length(swept) == 0) return(rep("default", nrow(df)))
  single_param <- length(swept) == 1
  parts <- list()
  if ("ct" %in% swept) {
    formatted <- format_number(df$ct)
    parts[["CT"]] <- if (single_param) formatted else paste0("CT=", formatted)
  }
  if ("hybridization_temperature" %in% swept) {
    temp_val <- as.character(df$hybridization_temperature)
    parts[["Temp"]] <- if (single_param) temp_val else paste0("Temp=", temp_val, "°C")
  }
  if ("capture_fraction" %in% swept) {
    cf_val <- as.character(df$capture_fraction)
    parts[["CF"]] <- if (single_param) cf_val else paste0("CF=", cf_val)
  }
  if ("num_sequences" %in% swept) {
    ns_val <- ifelse(df$num_sequences == 0, "all", format_number(df$num_sequences))
    parts[["NS"]] <- if (single_param) ns_val else paste0("NS=", ns_val)
  }
  do.call(paste, c(parts, sep = ", "))
}

sweep_df$combo_label <- build_combo_label(sweep_df, swept_params)

# Order combo labels numerically by underlying parameter values
if (length(swept_params) > 0) {
  sort_cols <- intersect(c("ct", "hybridization_temperature", "capture_fraction", "num_sequences"), swept_params)
  order_df <- sweep_df %>%
    distinct(combo_label, .keep_all = TRUE) %>%
    arrange(across(all_of(sort_cols)))
  sweep_df$combo_label <- factor(sweep_df$combo_label, levels = order_df$combo_label)
}

n_combos <- length(unique(sweep_df$combo_label))

# Legend title: full parameter name when only one type is swept
legend_title <- if (length(swept_params) == 1) {
  param_full_names[swept_params]
} else {
  "Parameters"
}

# Sample display for title
sample_display <- paste(sample_ids, collapse = ", ")
```

**Generated:** `r Sys.time()`

**Sample targets:** `r sample_display` | **Swept parameters:** `r if (length(swept_params) > 0) paste(swept_params, collapse = ", ") else "none (single run)"`

```{r params_fold, eval=has_params, results='asis'}
cat('<details><summary>Parameters</summary>\n\n')
```

```{r params_table, eval=has_params}
display_params <- params_df %>% dplyr::select(parameter, value)
knitr::kable(display_params, col.names = c("Parameter", "Value"), align = "ll")
```

```{r params_command, eval=has_params, results='asis'}
cmd_parts <- c("baitbench coverage-curve")
for (i in seq_len(nrow(params_df))) {
  flag <- params_df$flag[i]
  value <- params_df$value[i]
  if (value == "none" || flag == "(info)") next
  cmd_parts <- c(cmd_parts, sprintf("  %s %s", flag, value))
}
cat("```bash\n")
cat(paste(cmd_parts, collapse = " \\\n"), "\n")
cat("```\n")
```

```{r params_fold_end, eval=has_params, results='asis'}
cat('</details>\n\n')
```

---

## Coverage Depth Curves

Percentage of each target genome covered at increasing depth thresholds.
`r if (length(swept_params) > 0) "Each line represents a different parameter combination." else ""`

```{r depth-curves, fig.height=if (n_targets > 1) min(max(6, ceiling(n_targets / 2) * 4), 60) else 6}
# Determine faceting / color strategy.
# Rule: if num_sequences is swept, it is always the color axis; other params facet.
# Otherwise: facet by the param with fewest levels when there are >=10 combos.
use_facet_param <- FALSE
color_label <- "combo_label"

if ("num_sequences" %in% swept_params) {
  ns_order <- sweep_df %>% distinct(num_sequences) %>% arrange(num_sequences)
  sweep_df$ns_label <- factor(
    ifelse(sweep_df$num_sequences == 0, "all", format_number(sweep_df$num_sequences)),
    levels = ifelse(ns_order$num_sequences == 0, "all", format_number(ns_order$num_sequences))
  )
  color_label  <- "ns_label"
  legend_title <- "Num Sequences"

  remaining_swept <- setdiff(swept_params, "num_sequences")
  if (length(remaining_swept) > 0) {
    use_facet_param <- TRUE
    sweep_df$facet_label <- build_combo_label(sweep_df, remaining_swept)
    facet_sort_cols <- intersect(
      c("ct", "hybridization_temperature", "capture_fraction"), remaining_swept
    )
    facet_order_df <- sweep_df %>%
      distinct(facet_label, .keep_all = TRUE) %>%
      arrange(across(all_of(facet_sort_cols)))
    sweep_df$facet_label <- factor(sweep_df$facet_label, levels = facet_order_df$facet_label)
  }
} else if (n_combos >= 10 && length(swept_params) >= 2) {
  param_levels <- sapply(swept_params, function(p) length(unique(sweep_df[[p]])))
  facet_param  <- swept_params[which.min(param_levels)]
  use_facet_param <- TRUE

  remaining_swept <- setdiff(swept_params, facet_param)
  sweep_df$color_label <- build_combo_label(sweep_df, remaining_swept)
  color_sort_cols <- intersect(
    c("ct", "hybridization_temperature", "capture_fraction", "num_sequences"), remaining_swept
  )
  color_order_df <- sweep_df %>%
    distinct(color_label, .keep_all = TRUE) %>%
    arrange(across(all_of(color_sort_cols)))
  sweep_df$color_label <- factor(sweep_df$color_label, levels = color_order_df$color_label)
  legend_title <- if (length(remaining_swept) == 1) param_full_names[remaining_swept] else "Parameters"

  if (facet_param == "ct") {
    sweep_df$facet_label <- paste0("CT=", format_number(sweep_df$ct))
  } else if (facet_param == "hybridization_temperature") {
    sweep_df$facet_label <- paste0("Temp=", sweep_df$hybridization_temperature, "°C")
  } else if (facet_param == "capture_fraction") {
    sweep_df$facet_label <- paste0("CF=", sweep_df$capture_fraction)
  } else if (facet_param == "num_sequences") {
    sweep_df$facet_label <- ifelse(sweep_df$num_sequences == 0, "NS=all",
                                    paste0("NS=", format_number(sweep_df$num_sequences)))
  }
  facet_order_df <- sweep_df %>%
    distinct(facet_label, .keep_all = TRUE) %>%
    arrange(.data[[facet_param]])
  sweep_df$facet_label <- factor(sweep_df$facet_label, levels = facet_order_df$facet_label)
  color_label <- "color_label"
}

# Build plot
if (n_combos == 1) {
  p <- ggplot(sweep_df, aes(x = depth_threshold, y = pct_covered)) +
    geom_line(linewidth = 1, alpha = 0.85, color = "#440154") +
    labs(title = paste0("Coverage Depth Curves — ", sample_display))
} else {
  n_colors     <- length(unique(sweep_df[[color_label]]))
  combo_colors <- scales::viridis_pal(option = "plasma")(max(n_colors, 2))

  p <- ggplot(sweep_df, aes(x = depth_threshold, y = pct_covered,
                             color = .data[[color_label]],
                             group = combo_label)) +
    geom_line(linewidth = 1, alpha = 0.85) +
    scale_color_manual(values = combo_colors, name = legend_title) +
    labs(title = paste0("Coverage Depth Curves — ", sample_display))
}

p <- p +
  scale_x_log10(
    breaks = c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000),
    labels = function(x) ifelse(x >= 1000, paste0(x/1000, "k"), x)
  ) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 10)) +
  labs(x = "Depth of Coverage (X)", y = "% Genome Covered") +
  theme(legend.position = "right", plot.title = element_text(face = "bold"))

# Add faceting
if (use_facet_param && n_targets > 1) {
  p <- p + facet_grid(reference_id ~ facet_label, scales = "free_x")
} else if (use_facet_param) {
  p <- p + facet_wrap(~ facet_label, scales = "free_x")
} else if (n_targets > 1) {
  p <- p + facet_wrap(~ reference_id, ncol = 2, scales = "free_x")
}

print(p)
```

## Summary Table

Coverage at key depth thresholds for each parameter combination and sample target.

```{r summary-table}
key_thresholds <- c(1, 5, 10, 20, 50, 100)
available_thresholds <- key_thresholds[key_thresholds %in% sweep_df$depth_threshold]

if (length(available_thresholds) > 0) {
  summary_tbl <- sweep_df %>%
    filter(depth_threshold %in% available_thresholds) %>%
    mutate(threshold_label = paste0("≥", depth_threshold, "X"))

  # Build display columns based on swept params
  display_cols <- c()
  if ("ct" %in% swept_params) {
    display_cols <- c(display_cols, "ct")
  }
  if ("hybridization_temperature" %in% swept_params) {
    display_cols <- c(display_cols, "hybridization_temperature")
  }
  if ("capture_fraction" %in% swept_params) {
    display_cols <- c(display_cols, "capture_fraction")
  }
  if ("num_sequences" %in% swept_params) {
    summary_tbl <- summary_tbl %>%
      mutate(num_sequences_display = ifelse(num_sequences == 0, "all", as.character(num_sequences)))
    display_cols <- c(display_cols, "num_sequences_display")
  }

  id_cols <- c(display_cols, "reference_id")

  summary_wide <- summary_tbl %>%
    select(all_of(c(id_cols, "threshold_label", "pct_covered"))) %>%
    pivot_wider(names_from = threshold_label, values_from = pct_covered) %>%
    arrange(across(all_of(id_cols)))

  # Build column names
  col_names <- c()
  if ("ct" %in% swept_params) col_names <- c(col_names, "CT")
  if ("hybridization_temperature" %in% swept_params) col_names <- c(col_names, "Hyb Temp (°C)")
  if ("capture_fraction" %in% swept_params) col_names <- c(col_names, "Capture Fraction")
  if ("num_sequences" %in% swept_params) col_names <- c(col_names, "Num Sequences")
  col_names <- c(col_names, "Target")
  col_names <- c(col_names, names(summary_wide)[-(1:length(c(display_cols, "reference_id")))])

  if (nrow(summary_wide) > 20) {
    datatable(
      summary_wide,
      colnames = col_names,
      options = list(pageLength = 25, scrollX = TRUE),
      rownames = FALSE,
      filter = if (nrow(summary_wide) > 50) "top" else "none"
    ) %>% formatRound(names(summary_wide)[-(1:length(id_cols))], 1)
  } else {
    knitr::kable(
      summary_wide,
      digits = 1,
      col.names = col_names,
      align = c(rep("l", length(id_cols)), rep("r", ncol(summary_wide) - length(id_cols)))
    )
  }
} else {
  cat("No data available at standard depth thresholds.")
}
```

---

*Generated by [BaitBench](https://github.com/CrossLab/BaitBench) coverage-curve*
