---
title: "BaitBench Coverage Curve Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  sweep_file: ""
  sample_ids: !r c()
  swept_params: !r c()
---

<style>
.main-container { max-width: 100%; margin-left: 1em; margin-right: 1em; }
table { width: 100%; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 6, out.width = "100%")

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
})

theme_set(theme_minimal(base_size = 14))

# Read data
sweep_df <- read.delim(params$sweep_file, stringsAsFactors = FALSE)

sample_ids <- params$sample_ids
swept_params <- params$swept_params
n_targets <- length(unique(sweep_df$reference_id))

# Detect swept params from data if not explicitly provided
if (length(swept_params) == 0 || (length(swept_params) == 1 && swept_params[1] == "")) {
  swept_params <- character(0)
  if (length(unique(sweep_df$ct)) > 1) swept_params <- c(swept_params, "ct")
  if (length(unique(sweep_df$fold_enrichment)) > 1) swept_params <- c(swept_params, "fold_enrichment")
  if (length(unique(sweep_df$num_sequences)) > 1) swept_params <- c(swept_params, "num_sequences")
}

# Build combo label from swept params
build_combo_label <- function(df, swept) {
  if (length(swept) == 0) return(rep("default", nrow(df)))
  parts <- list()
  if ("ct" %in% swept) parts[["CT"]] <- paste0("CT=", df$ct)
  if ("fold_enrichment" %in% swept) {
    parts[["FE"]] <- ifelse(df$fold_enrichment == 0, "FE=none",
                            paste0("FE=", df$fold_enrichment))
  }
  if ("num_sequences" %in% swept) {
    parts[["NS"]] <- ifelse(df$num_sequences == 0, "NS=all",
                            paste0("NS=", df$num_sequences))
  }
  do.call(paste, c(parts, sep = ", "))
}

sweep_df$combo_label <- build_combo_label(sweep_df, swept_params)
n_combos <- length(unique(sweep_df$combo_label))

# Sample display for title
sample_display <- paste(sample_ids, collapse = ", ")
```

**Generated:** `r Sys.time()`

**Sample targets:** `r sample_display` | **Swept parameters:** `r if (length(swept_params) > 0) paste(swept_params, collapse = ", ") else "none (single run)"`

---

## Coverage Depth Curves

Percentage of each target genome covered at increasing depth thresholds.
`r if (length(swept_params) > 0) "Each line represents a different parameter combination." else ""`

```{r depth-curves, fig.height=if (n_targets > 1) max(6, ceiling(n_targets / 2) * 4) else 6}
# Determine faceting strategy
use_facet_param <- FALSE
facet_param <- NULL
color_label <- "combo_label"

if (n_combos >= 10 && length(swept_params) >= 2) {
  # Find swept param with fewest levels for faceting
  param_levels <- sapply(swept_params, function(p) {
    length(unique(sweep_df[[p]]))
  })
  facet_param <- swept_params[which.min(param_levels)]
  use_facet_param <- TRUE

  # Build color label from remaining swept params
  remaining_swept <- setdiff(swept_params, facet_param)
  sweep_df$color_label <- build_combo_label(sweep_df, remaining_swept)

  # Build facet label
  if (facet_param == "ct") {
    sweep_df$facet_label <- paste0("CT=", sweep_df$ct)
  } else if (facet_param == "fold_enrichment") {
    sweep_df$facet_label <- ifelse(sweep_df$fold_enrichment == 0, "FE=none",
                                    paste0("FE=", sweep_df$fold_enrichment))
  } else if (facet_param == "num_sequences") {
    sweep_df$facet_label <- ifelse(sweep_df$num_sequences == 0, "NS=all",
                                    paste0("NS=", sweep_df$num_sequences))
  }
  color_label <- "color_label"
}

# Build plot
if (n_combos == 1) {
  # Single combo — no color legend
  p <- ggplot(sweep_df, aes(x = depth_threshold, y = pct_covered)) +
    geom_line(linewidth = 1, alpha = 0.85, color = "#440154") +
    labs(title = paste0("Coverage Depth Curves \u2014 ", sample_display))
} else if (!use_facet_param) {
  # <10 combos — all on one plot, color by combo
  n_colors <- n_combos
  combo_colors <- scales::viridis_pal(option = "plasma")(n_colors)

  p <- ggplot(sweep_df, aes(x = depth_threshold, y = pct_covered,
                             color = combo_label, group = combo_label)) +
    geom_line(linewidth = 1, alpha = 0.85) +
    scale_color_manual(values = combo_colors, name = "Parameters") +
    labs(title = paste0("Coverage Depth Curves \u2014 ", sample_display))
} else {
  # >=10 combos — facet by param with fewest levels
  n_colors <- length(unique(sweep_df[[color_label]]))
  combo_colors <- scales::viridis_pal(option = "plasma")(max(n_colors, 2))

  p <- ggplot(sweep_df, aes(x = depth_threshold, y = pct_covered,
                             color = .data[[color_label]],
                             group = combo_label)) +
    geom_line(linewidth = 1, alpha = 0.85) +
    scale_color_manual(values = combo_colors, name = "Parameters") +
    labs(title = paste0("Coverage Depth Curves \u2014 ", sample_display))
}

p <- p +
  scale_x_log10(
    breaks = c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000),
    labels = function(x) ifelse(x >= 1000, paste0(x/1000, "k"), x)
  ) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 10)) +
  labs(
    x = "Depth of Coverage (X)",
    y = "% Genome Covered"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )

# Add faceting
if (use_facet_param && n_targets > 1) {
  p <- p + facet_grid(reference_id ~ facet_label, scales = "free_x")
} else if (use_facet_param) {
  p <- p + facet_wrap(~ facet_label, scales = "free_x")
} else if (n_targets > 1) {
  p <- p + facet_wrap(~ reference_id, ncol = 2, scales = "free_x")
}

print(p)
```

## Summary Table

Coverage at key depth thresholds for each parameter combination and sample target.

```{r summary-table}
key_thresholds <- c(1, 5, 10, 20, 50, 100)
available_thresholds <- key_thresholds[key_thresholds %in% sweep_df$depth_threshold]

if (length(available_thresholds) > 0) {
  summary_tbl <- sweep_df %>%
    filter(depth_threshold %in% available_thresholds) %>%
    mutate(threshold_label = paste0("\u2265", depth_threshold, "X"))

  # Build display columns based on swept params
  display_cols <- c()
  if ("ct" %in% swept_params) {
    display_cols <- c(display_cols, "ct")
  }
  if ("fold_enrichment" %in% swept_params) {
    summary_tbl <- summary_tbl %>%
      mutate(fold_enrichment_display = ifelse(fold_enrichment == 0, "none", as.character(fold_enrichment)))
    display_cols <- c(display_cols, "fold_enrichment_display")
  }
  if ("num_sequences" %in% swept_params) {
    summary_tbl <- summary_tbl %>%
      mutate(num_sequences_display = ifelse(num_sequences == 0, "all", as.character(num_sequences)))
    display_cols <- c(display_cols, "num_sequences_display")
  }

  id_cols <- c(display_cols, "reference_id")

  summary_wide <- summary_tbl %>%
    select(all_of(c(id_cols, "threshold_label", "pct_covered"))) %>%
    pivot_wider(names_from = threshold_label, values_from = pct_covered) %>%
    arrange(across(all_of(id_cols)))

  # Build column names
  col_names <- c()
  if ("ct" %in% swept_params) col_names <- c(col_names, "CT")
  if ("fold_enrichment" %in% swept_params) col_names <- c(col_names, "Fold Enrichment")
  if ("num_sequences" %in% swept_params) col_names <- c(col_names, "Num Sequences")
  col_names <- c(col_names, "Target")
  col_names <- c(col_names, names(summary_wide)[-(1:length(c(display_cols, "reference_id")))])

  knitr::kable(
    summary_wide,
    digits = 1,
    col.names = col_names,
    align = c(rep("l", length(id_cols)), rep("r", ncol(summary_wide) - length(id_cols)))
  )
} else {
  cat("No data available at standard depth thresholds.")
}
```

---

*Generated by [BaitBench](https://github.com/CrossLab/BaitBench) coverage-curve*
