---
title: "BaitBench CT Sweep Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  sweep_file: ""
  sample_ids: !r c()
---

<style>
.main-container { max-width: 100%; margin-left: 1em; margin-right: 1em; }
table { width: 100%; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 6, out.width = "100%")

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
})

theme_set(theme_minimal(base_size = 14))

# Read data
sweep_df <- read.delim(params$sweep_file, stringsAsFactors = FALSE)
sweep_df$ct <- as.factor(sweep_df$ct)

sample_ids <- params$sample_ids
n_targets <- length(unique(sweep_df$reference_id))
```

**Generated:** `r Sys.time()`

**Sample targets:** `r paste(sample_ids, collapse = ", ")` | **CT values:** `r paste(levels(sweep_df$ct), collapse = ", ")`

---

## Coverage Depth Curves

Percentage of each target genome covered at increasing depth thresholds. Each line represents a different CT value (lower CT = more target material).

```{r depth-curves, fig.height=if (n_targets > 1) max(6, ceiling(n_targets / 2) * 4) else 6}
ct_colors <- scales::viridis_pal(option = "plasma")(length(levels(sweep_df$ct)))

p <- ggplot(sweep_df, aes(x = depth_threshold, y = pct_covered,
                           color = ct, group = ct)) +
  geom_line(linewidth = 1, alpha = 0.85) +
  scale_x_log10(
    breaks = c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000),
    labels = function(x) ifelse(x >= 1000, paste0(x/1000, "k"), x)
  ) +
  scale_color_manual(values = ct_colors, name = "CT Value") +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 10)) +
  labs(
    x = "Depth of Coverage (X)",
    y = "% Genome Covered",
    title = "Coverage Depth Curves by CT Value"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )

if (n_targets > 1) {
  p <- p + facet_wrap(~ reference_id, ncol = 2, scales = "free_x")
}

print(p)
```

## Summary Table

Coverage at key depth thresholds for each CT value and sample target.

```{r summary-table}
key_thresholds <- c(1, 5, 10, 20, 50, 100)
available_thresholds <- key_thresholds[key_thresholds %in% sweep_df$depth_threshold]

if (length(available_thresholds) > 0) {
  summary_tbl <- sweep_df %>%
    filter(depth_threshold %in% available_thresholds) %>%
    mutate(threshold_label = paste0("\u2265", depth_threshold, "X")) %>%
    select(ct, reference_id, threshold_label, pct_covered) %>%
    pivot_wider(names_from = threshold_label, values_from = pct_covered) %>%
    arrange(reference_id, ct)

  knitr::kable(
    summary_tbl,
    digits = 1,
    col.names = c("CT", "Target", names(summary_tbl)[-(1:2)]),
    align = c("l", "l", rep("r", ncol(summary_tbl) - 2))
  )
} else {
  cat("No data available at standard depth thresholds.")
}
```

---

*Generated by [BaitBench](https://github.com/CrossLab/BaitBench) ct-sweep*
