---
title: "BaitBench Target Panel Discriminability Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  discriminability_file: ""
  matrix_file: ""
  similarity_file: ""
  params_file: ""
---

<style>
.main-container { max-width: 100%; margin-left: 1em; margin-right: 1em; }
.dataTables_wrapper { width: 100%; }
table { width: 100%; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 6, out.width = "100%")

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  library(plotly)
  library(DT)
})

theme_set(theme_minimal(base_size = 14))

disc <- read.delim(params$discriminability_file, stringsAsFactors = FALSE)
matrix_df <- read.delim(params$matrix_file, stringsAsFactors = FALSE, check.names = FALSE)

has_similarity <- params$similarity_file != "" && file.exists(params$similarity_file)
if (has_similarity) {
  sim <- read.delim(params$similarity_file, stringsAsFactors = FALSE)
}

has_params <- params$params_file != "" && file.exists(params$params_file)
if (has_params) {
  params_df <- read.delim(params$params_file, stringsAsFactors = FALSE)
}

n_species <- nrow(disc)
n_targets <- sum(disc$total_targets)
n_unique_species <- sum(disc$discriminability_score >= 1.0)
n_partial_species <- sum(disc$discriminability_score > 0 & disc$discriminability_score < 1.0)
n_zero_species <- sum(disc$discriminability_score == 0)
```

## Panel Summary

```{r summary_table, results='asis'}
info <- data.frame(
  Metric = c(
    "Total species",
    "Total targets",
    "Species with all unique targets",
    "Species with some unique targets",
    "Species with NO unique targets"
  ),
  Value = c(
    n_species,
    n_targets,
    paste0(n_unique_species, " / ", n_species, " (", round(100 * n_unique_species / max(n_species, 1), 1), "%)"),
    paste0(n_partial_species, " / ", n_species, " (", round(100 * n_partial_species / max(n_species, 1), 1), "%)"),
    paste0(n_zero_species, " / ", n_species, " (", round(100 * n_zero_species / max(n_species, 1), 1), "%)")
  ),
  stringsAsFactors = FALSE
)

knitr::kable(info, col.names = c("Metric", "Value"), align = "lr")
```

`r if (n_zero_species > 0) paste0("**Warning:** ", n_zero_species, " species have no unique targets and cannot be reliably distinguished from other species. Consider adding more species-specific targets to the panel.")`

## Species Discriminability

```{r discriminability_plot, fig.height=if (n_species <= 50) max(4, n_species * 0.3) else 5}
disc_plot <- disc %>%
  mutate(
    species_id = factor(species_id, levels = species_id[order(discriminability_score)]),
    tier = case_when(
      discriminability_score >= 0.5 ~ "High (>= 50%)",
      discriminability_score > 0    ~ "Low (> 0%)",
      TRUE                          ~ "None (0%)"
    ),
    tier = factor(tier, levels = c("High (>= 50%)", "Low (> 0%)", "None (0%)"))
  )

tier_colors <- c("High (>= 50%)" = "#1a7a42", "Low (> 0%)" = "#5b9bd5", "None (0%)" = "#e74c3c")

if (n_species <= 50) {
  ggplot(disc_plot, aes(x = species_id, y = discriminability_score, fill = tier)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = paste0(unique_targets, "/", total_targets)),
              hjust = -0.1, size = 3) +
    scale_fill_manual(values = tier_colors, name = "Discriminability") +
    scale_y_continuous(limits = c(0, 1.15), labels = scales::percent_format()) +
    coord_flip() +
    labs(x = NULL, y = "Discriminability Score (unique / total targets)",
         title = "Species Discriminability") +
    theme(legend.position = "bottom")
} else {
  # For large panels, show distribution instead
  ggplot(disc_plot, aes(x = discriminability_score, fill = tier)) +
    geom_histogram(bins = 20, boundary = 0) +
    scale_fill_manual(values = tier_colors, name = "Discriminability") +
    scale_x_continuous(labels = scales::percent_format()) +
    labs(x = "Discriminability Score", y = "Number of Species",
         title = "Distribution of Species Discriminability Scores") +
    theme(legend.position = "bottom")
}
```

## Target Composition

```{r target_composition, fig.height=if (n_species <= 50) max(4, n_species * 0.3) else 5}
comp <- disc %>%
  select(species_id, unique_targets, shared_targets) %>%
  pivot_longer(cols = c(unique_targets, shared_targets),
               names_to = "type", values_to = "count") %>%
  mutate(
    species_id = factor(species_id, levels = disc$species_id[order(disc$discriminability_score)]),
    type = factor(type, levels = c("shared_targets", "unique_targets"),
                  labels = c("Shared", "Unique"))
  )

comp_colors <- c("Unique" = "#1a7a42", "Shared" = "#5b9bd5")

if (n_species <= 50) {
  ggplot(comp, aes(x = species_id, y = count, fill = type)) +
    geom_col(position = "identity", width = 0.7, alpha = 0.6) +
    scale_fill_manual(values = comp_colors, name = "Target Type") +
    coord_flip() +
    labs(x = NULL, y = "Number of Targets",
         title = "Target Composition per Species (Unique vs Shared)") +
    theme(legend.position = "bottom")
} else {
  ggplot(comp, aes(x = count, fill = type)) +
    geom_density(alpha = 0.5, adjust = 1.2) +
    scale_fill_manual(values = comp_colors, name = "Target Type") +
    labs(x = "Number of Targets", y = "Density",
         title = "Distribution of Unique vs Shared Targets per Species") +
    theme(legend.position = "bottom")
}
```

## Species Confusion Matrix

```{r confusion_matrix_setup, include=FALSE}
# The matrix_df has species as first column, rest are species columns
species_names <- matrix_df[[1]]
mat <- as.matrix(matrix_df[, -1])
rownames(mat) <- species_names
heatmap_height <- max(500, n_species * 12)
```

```{r confusion_matrix_heatmap, eval=(n_species <= 100), out.height=paste0(heatmap_height, "px")}
# Build hover text matrix: show species names and count for non-zero cells
hover_text <- matrix("", nrow = n_species, ncol = n_species)
for (i in seq_len(n_species)) {
  for (j in seq_len(n_species)) {
    if (mat[i, j] > 0) {
      hover_text[i, j] <- paste0(species_names[i], " \u2194 ", species_names[j],
                                  "\nShared targets: ", mat[i, j])
    }
  }
}

# Replace zeros with NA for cleaner color scale (white for zero)
mat_plot <- mat
mat_plot[mat_plot == 0] <- NA

plot_ly(
  x = species_names, y = species_names, z = mat_plot,
  text = hover_text, hoverinfo = "text",
  type = "heatmap",
  colorscale = list(c(0, "#fee0d2"), c(0.5, "#fc9272"), c(1, "#de2d26")),
  colorbar = list(title = "Shared\nTargets"),
  zmin = 1
) %>%
  layout(
    title = "Cross-Species Target Similarity (shared target count)",
    xaxis = list(tickangle = -90, tickfont = list(size = if (n_species > 50) 7 else 10)),
    yaxis = list(tickfont = list(size = if (n_species > 50) 7 else 10), autorange = "reversed"),
    height = heatmap_height,
    margin = list(b = 120)
  )
```

```{r confusion_matrix_text, eval=(n_species > 100), results='asis'}
off_diag <- mat[upper.tri(mat)]
cat("**Panel too large for interactive heatmap (", n_species, " species).** Summary of off-diagonal confusion counts:\n\n")
cat("- Species pairs with shared targets:", sum(off_diag > 0), "/", length(off_diag), "\n")
cat("- Max shared targets between any pair:", max(off_diag), "\n")
cat("- Mean shared targets (non-zero pairs):", round(mean(off_diag[off_diag > 0]), 1), "\n")
```

## Discriminability Table

```{r disc_table}
disc_display <- disc %>%
  mutate(discriminability_score = round(discriminability_score, 3)) %>%
  select(species_id, total_targets, unique_targets, shared_targets,
         discriminability_score, confusable_species)

if (n_species <= 20) {
  knitr::kable(disc_display,
               col.names = c("Species", "Total Targets", "Unique", "Shared",
                              "Discriminability", "Confusable With"),
               align = "lrrrrl")
} else {
  datatable(disc_display,
            colnames = c("Species", "Total Targets", "Unique", "Shared",
                         "Discriminability", "Confusable With"),
            options = list(pageLength = 25, scrollX = TRUE),
            rownames = FALSE) %>%
    formatRound("discriminability_score", digits = 3)
}
```

```{r similarity_table, eval=has_similarity, results='asis'}
cat("\n## Target Similarity Pairs\n\n")

sim_display <- sim %>%
  mutate(identity_pct = round(identity_pct, 1)) %>%
  arrange(desc(identity_pct))

if (nrow(sim_display) <= 20) {
  knitr::kable(sim_display,
               col.names = c("Target A", "Target B", "Identity %",
                              "Matching Bases", "Length A", "Length B"),
               align = "llrrrr")
} else {
  datatable(sim_display,
            colnames = c("Target A", "Target B", "Identity %",
                         "Matching Bases", "Length A", "Length B"),
            options = list(pageLength = 25, scrollX = TRUE),
            rownames = FALSE) %>%
    formatRound("identity_pct", digits = 1)
}
```

```{r params_section, eval=has_params, results='asis'}
cat("\n\n<details><summary>Parameters</summary>\n\n")

knitr::kable(params_df, col.names = c("Parameter", "Flag", "Value"), align = "lll")

# Reconstruct CLI command
cmd_parts <- c("baitbench panel-qc")
for (i in seq_len(nrow(params_df))) {
  flag <- params_df$flag[i]
  val <- params_df$value[i]
  if (!is.na(val) && val != "" && val != "none" && val != "false") {
    if (val == "true") {
      cmd_parts <- c(cmd_parts, flag)
    } else {
      cmd_parts <- c(cmd_parts, paste(flag, val))
    }
  }
}

cat("\n**Command:**\n```\n", paste(cmd_parts, collapse = " \\\n  "), "\n```\n")
cat("\n</details>\n")
```
