---
title: "BaitBench Probe Coverage Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  summary_file: ""
  depth_file: ""
  multi_mapping_file: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 5)

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
})

theme_set(theme_minimal(base_size = 14))

# Read data
summary_df <- read.delim(params$summary_file, stringsAsFactors = FALSE)
depth_df <- read.delim(params$depth_file, stringsAsFactors = FALSE)

# Read multi-mapping probes if available
has_multi <- nchar(params$multi_mapping_file) > 0 && file.exists(params$multi_mapping_file)
if (has_multi) {
  multi_df <- read.delim(params$multi_mapping_file, stringsAsFactors = FALSE)
} else {
  multi_df <- data.frame(probe_id = character(), num_targets = integer(), targets = character())
}
```

**Generated:** `r Sys.time()`

**Targets analyzed:** `r nrow(summary_df)` | **Fully tiled (100%):** `r sum(summary_df$pct_covered_1x >= 100)` | **Well tiled (>=90%):** `r sum(summary_df$pct_covered_1x >= 90)` | **Multi-mapping probes:** `r nrow(multi_df)`

---

## Summary Table

```{r summary-table}
summary_df %>%
  arrange(pct_covered_1x) %>%
  mutate(
    ref_length = format(ref_length, big.mark = ","),
    covered_bases = format(covered_bases, big.mark = ","),
    pct_covered_1x = sprintf("%.1f%%", pct_covered_1x),
    mean_depth = sprintf("%.1f", mean_depth),
    median_depth = sprintf("%.1f", median_depth),
    pct_covered_2x = sprintf("%.1f%%", pct_covered_2x),
    pct_covered_5x = sprintf("%.1f%%", pct_covered_5x),
    pct_covered_10x = sprintf("%.1f%%", pct_covered_10x),
    max_gap_length = format(max_gap_length, big.mark = ","),
    pct_near_probe = sprintf("%.1f%%", pct_near_probe)
  ) %>%
  knitr::kable(
    col.names = c("Reference", "Length", "Covered", "%\u22651X", "Mean Depth",
                   "Median Depth", "%\u22652X", "%\u22655X", "%\u226510X",
                   "Max Gap", "Gaps", "% Near Probe"),
    align = "lrrrrrrrrrrr"
  )
```

## Coverage Breadth

```{r coverage-breadth, fig.height=max(3, nrow(summary_df) * 0.4)}
summary_df <- summary_df %>%
  mutate(
    coverage_class = case_when(
      pct_covered_1x >= 95 ~ "Excellent (>=95%)",
      pct_covered_1x >= 80 ~ "Good (>=80%)",
      TRUE ~ "Poor (<80%)"
    ),
    coverage_class = factor(coverage_class,
                            levels = c("Excellent (>=95%)", "Good (>=80%)", "Poor (<80%)"))
  )

ggplot(summary_df, aes(x = reorder(reference_id, pct_covered_1x),
                        y = pct_covered_1x, fill = coverage_class)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.3) +
  coord_flip() +
  scale_fill_manual(values = c("Excellent (>=95%)" = "#27ae60",
                                "Good (>=80%)" = "#f39c12",
                                "Poor (<80%)" = "#e74c3c")) +
  labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
       title = "Probe Tiling Breadth per Target") +
  theme(legend.position = "bottom")
```

## Tiered Coverage

```{r tiered-coverage, fig.height=max(3, nrow(summary_df) * 0.5)}
tiered <- summary_df %>%
  select(reference_id, pct_covered_1x, pct_covered_2x, pct_covered_5x, pct_covered_10x) %>%
  pivot_longer(cols = starts_with("pct_covered_"),
               names_to = "threshold", values_to = "pct") %>%
  mutate(threshold = case_when(
    threshold == "pct_covered_1x"  ~ ">=1X",
    threshold == "pct_covered_2x"  ~ ">=2X",
    threshold == "pct_covered_5x"  ~ ">=5X",
    threshold == "pct_covered_10x" ~ ">=10X"
  ),
  threshold = factor(threshold, levels = c(">=1X", ">=2X", ">=5X", ">=10X")))

ggplot(tiered, aes(x = reorder(reference_id, pct), y = pct, fill = threshold)) +
  geom_col(position = "dodge", width = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c(">=1X" = "#3498db", ">=2X" = "#2ecc71",
                                ">=5X" = "#f39c12", ">=10X" = "#e74c3c")) +
  labs(x = NULL, y = "% Bases Covered", fill = "Depth Threshold",
       title = "Tiered Probe Coverage per Target") +
  theme(legend.position = "bottom")
```

## Gap Analysis

```{r gap-analysis, fig.height=max(3, nrow(summary_df) * 0.4)}
gap_data <- summary_df %>% filter(num_gaps > 0)

if (nrow(gap_data) > 0) {
  ggplot(gap_data, aes(x = reorder(reference_id, max_gap_length), y = max_gap_length)) +
    geom_segment(aes(xend = reference_id, y = 0, yend = max_gap_length),
                 color = "#e74c3c", linewidth = 1) +
    geom_point(size = 4, color = "#e74c3c") +
    geom_text(aes(label = sprintf("%s bp (%d gaps)",
                                   format(max_gap_length, big.mark = ","), num_gaps)),
              hjust = -0.1, size = 3.5) +
    coord_flip() +
    labs(x = NULL, y = "Largest Gap (bp)",
         title = "Largest Uncovered Stretch per Target",
         subtitle = "Targets with no gaps are omitted") +
    expand_limits(y = max(gap_data$max_gap_length) * 1.3)
} else {
  cat("All targets are fully tiled with no gaps.")
}
```

## Pangenome Probe Depth

Targets concatenated end-to-end, sorted by name. Vertical grey lines mark target boundaries.

```{r pangenome, fig.height=4, fig.width=10}
if (nrow(depth_df) > 0) {
  # Build cumulative offsets for each target
  target_order <- summary_df %>% arrange(reference_id)
  offsets <- target_order %>%
    mutate(offset = cumsum(lag(ref_length, default = 0)))

  # Join offsets and compute pangenome position
  pan_df <- depth_df %>%
    inner_join(offsets %>% select(reference_id, offset), by = "reference_id") %>%
    mutate(pan_position = position + offset)

  # Downsample for rendering speed
  pan_df <- pan_df %>%
    mutate(bin = ceiling(row_number() / max(1, ceiling(n() / 2000)))) %>%
    group_by(bin) %>%
    summarise(pan_position = mean(pan_position), depth = mean(depth),
              reference_id = first(reference_id), .groups = "drop")

  # Target boundary positions for vertical lines
  boundaries <- offsets %>%
    mutate(boundary = offset + ref_length) %>%
    filter(row_number() < n())  # no line after last target

  # Midpoints for target labels
  labels_df <- offsets %>%
    mutate(mid = offset + ref_length / 2)

  ggplot(pan_df, aes(x = pan_position, y = depth)) +
    geom_area(fill = "#3498db", alpha = 0.4) +
    geom_line(color = "#2c3e50", linewidth = 0.2) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.3) +
    geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.3) +
    { if (nrow(boundaries) > 0 && nrow(target_order) <= 12)
        geom_vline(data = boundaries, aes(xintercept = boundary),
                   color = "grey50", linetype = "solid", linewidth = 0.3) } +
    labs(x = "Pangenome Position", y = "Probe Depth",
         title = "Probe Depth Across Concatenated Targets",
         subtitle = sprintf("%d targets | Dashed lines: 1X (green), 5X (orange)",
                            nrow(target_order))) +
    theme(axis.text.x = element_text(size = 7))
}
```

## Probe Depth Profiles {.tabset}

```{r depth-overview, fig.height=max(3, ceiling(min(length(unique(depth_df$reference_id)), 12) / 3) * 2.5), fig.width=10}
if (nrow(depth_df) > 0) {
  # Limit faceted overview to 12 targets (worst-covered first)
  worst_refs <- summary_df %>%
    arrange(pct_covered_1x) %>%
    head(12) %>%
    pull(reference_id)

  overview_data <- depth_df %>% filter(reference_id %in% worst_refs)

  if (nrow(summary_df) > 12) {
    overview_subtitle <- sprintf("Showing 12 worst-covered of %d targets | Dashed lines: 1X (green), 5X (orange)",
                                  nrow(summary_df))
  } else {
    overview_subtitle <- "Dashed lines: 1X (green), 5X (orange)"
  }

  # Downsample for rendering speed
  plot_data <- overview_data %>%
    group_by(reference_id) %>%
    mutate(
      bin = ceiling(row_number() / max(1, ceiling(n() / 500)))
    ) %>%
    group_by(reference_id, bin) %>%
    summarise(position = mean(position), depth = mean(depth), .groups = "drop")

  ggplot(plot_data, aes(x = position, y = depth)) +
    geom_area(fill = "#3498db", alpha = 0.4) +
    geom_line(color = "#2c3e50", linewidth = 0.3) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.3) +
    geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.3) +
    facet_wrap(~reference_id, scales = "free", ncol = 3) +
    labs(x = "Position", y = "Probe Depth",
         title = "Probe Depth Across Targets",
         subtitle = overview_subtitle) +
    theme(strip.text = element_text(size = 8),
          axis.text = element_text(size = 6))
}
```

```{r depth-detail, results='asis', fig.width=8, fig.height=3}
if (nrow(depth_df) > 0 && nrow(summary_df) > 0) {
  # Show individual detail plots for the 10 worst-covered targets only
  ordered <- summary_df %>% arrange(pct_covered_1x)
  n_detail <- min(10, nrow(ordered))

  if (nrow(ordered) > 10) {
    cat(sprintf('\n\n*Showing detail for the %d worst-covered targets (of %d total).*\n\n',
                n_detail, nrow(ordered)))
  }

  for (i in seq_len(n_detail)) {
    ref_id <- ordered$reference_id[i]
    ref_data <- depth_df %>% filter(reference_id == ref_id)

    if (nrow(ref_data) == 0) next

    s <- ordered[i, ]
    cat(sprintf('\n\n<details><summary><strong>%s</strong> &mdash; %.1f%% covered, %.1fX mean, %s bp, %d gaps (max %s bp)</summary>\n\n',
                ref_id, s$pct_covered_1x, s$mean_depth,
                format(s$ref_length, big.mark = ","),
                s$num_gaps, format(s$max_gap_length, big.mark = ",")))

    # Identify zero-depth positions for gap rug marks
    zero_pos <- ref_data %>% filter(depth == 0)

    p <- ggplot(ref_data, aes(x = position, y = depth)) +
      geom_area(fill = "#3498db", alpha = 0.3) +
      geom_line(color = "#2c3e50", linewidth = 0.4) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.5) +
      geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.5)

    if (nrow(zero_pos) > 0) {
      p <- p + geom_rug(data = zero_pos, aes(x = position), color = "darkgreen",
                        alpha = 0.3, sides = "b", inherit.aes = FALSE)
    }

    p <- p +
      annotate("text", x = Inf, y = 1, label = "1X", hjust = 1.2, vjust = -0.5,
               color = "#27ae60", size = 3) +
      annotate("text", x = Inf, y = 5, label = "5X", hjust = 1.2, vjust = -0.5,
               color = "#f39c12", size = 3) +
      labs(x = "Genome Position", y = "Probe Depth",
           title = sprintf("Probe Depth: %s", ref_id),
           subtitle = sprintf("%.1f%% covered | %.1fX mean | %.1f median | %s bp",
                              s$pct_covered_1x, s$mean_depth, s$median_depth,
                              format(s$ref_length, big.mark = ",")),
           caption = if (nrow(zero_pos) > 0) "Rug: uncovered positions" else NULL) +
      theme_minimal(base_size = 12)

    print(p)
    cat('\n\n</details>\n\n')
  }
}
```

## Proximity Coverage

```{r proximity, fig.height=max(3, nrow(summary_df) * 0.4)}
ggplot(summary_df, aes(x = reorder(reference_id, pct_near_probe), y = pct_near_probe)) +
  geom_col(fill = "#8e44ad", width = 0.7, alpha = 0.8) +
  geom_col(aes(y = pct_covered_1x), fill = "#3498db", width = 0.7, alpha = 0.5) +
  coord_flip() +
  labs(x = NULL, y = "% Target Bases",
       title = "Direct Coverage vs Proximity Coverage",
       subtitle = "Blue: directly covered | Purple: within pull-down zone of a probe") +
  theme(legend.position = "bottom")
```

## Multi-Mapping Probes

`r nrow(multi_df)` probes map to more than one target sequence. See `multi_mapping_probes.tsv` for the full list.

```{r multi-mapping, eval=(nrow(multi_df) > 0)}
multi_df %>%
  arrange(desc(num_targets)) %>%
  head(20) %>%
  knitr::kable(
    col.names = c("Probe ID", "# Targets", "Targets"),
    align = "lrl",
    caption = if (nrow(multi_df) > 20)
      sprintf("Showing top 20 of %d multi-mapping probes", nrow(multi_df))
    else NULL
  )
```

---

*Generated by [BaitBench](https://github.com/CrossLab/BaitBench) probe-coverage*
