---
title: "BaitBench Probe Coverage Report"
output:
  html_document:
    self_contained: true
    theme: flatly
    toc: true
    toc_float: true
params:
  summary_file: ""
  depth_file: ""
  multi_mapping_file: ""
  proximity: 50
---

<style>
.main-container { max-width: 100%; margin-left: 1em; margin-right: 1em; }
.dataTables_wrapper { width: 100%; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 5)

suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
  library(tidyr)
  library(DT)
})

theme_set(theme_minimal(base_size = 14))

# Read data
summary_df <- read.delim(params$summary_file, stringsAsFactors = FALSE)
depth_df <- read.delim(params$depth_file, stringsAsFactors = FALSE)

# Read multi-mapping probes if available
has_multi <- nchar(params$multi_mapping_file) > 0 && file.exists(params$multi_mapping_file)
if (has_multi) {
  multi_df <- read.delim(params$multi_mapping_file, stringsAsFactors = FALSE)
} else {
  multi_df <- data.frame(probe_id = character(), num_targets = integer(), targets = character())
}

# Target count drives layout decisions
n_targets <- nrow(summary_df)
is_small  <- n_targets <= 20
is_medium <- n_targets > 20 & n_targets <= 100
is_large  <- n_targets > 100

# Proximity parameter for labels
proximity_bp <- params$proximity
```

**Generated:** `r Sys.time()`

**Targets analyzed:** `r n_targets` | **Fully tiled (100%):** `r sum(summary_df$pct_covered_1x >= 100)` | **Well tiled (>=90%):** `r sum(summary_df$pct_covered_1x >= 90)` | **Multi-mapping probes:** `r nrow(multi_df)` | **Proximity:** `r proximity_bp` bp

---

## Summary Table

```{r summary-table}
display_df <- summary_df %>%
  arrange(pct_covered_1x) %>%
  mutate(
    ref_length_fmt = format(ref_length, big.mark = ","),
    covered_bases_fmt = format(covered_bases, big.mark = ","),
    pct_covered_1x_fmt = sprintf("%.1f%%", pct_covered_1x),
    mean_depth_fmt = sprintf("%.1f", mean_depth),
    median_depth_fmt = sprintf("%.1f", median_depth),
    pct_covered_2x_fmt = sprintf("%.1f%%", pct_covered_2x),
    pct_covered_5x_fmt = sprintf("%.1f%%", pct_covered_5x),
    pct_covered_10x_fmt = sprintf("%.1f%%", pct_covered_10x),
    max_gap_length_fmt = format(max_gap_length, big.mark = ","),
    pct_near_probe_fmt = sprintf("%.1f%%", pct_near_probe)
  )

if (is_small) {
  display_df %>%
    select(reference_id, ref_length_fmt, covered_bases_fmt, pct_covered_1x_fmt,
           mean_depth_fmt, median_depth_fmt, pct_covered_2x_fmt, pct_covered_5x_fmt,
           pct_covered_10x_fmt, max_gap_length_fmt, num_gaps, pct_near_probe_fmt) %>%
    knitr::kable(
      col.names = c("Reference", "Length", "Covered", "%\u22651X", "Mean Depth",
                     "Median Depth", "%\u22652X", "%\u22655X", "%\u226510X",
                     "Max Gap", "Gaps", "% Near Probe"),
      align = "lrrrrrrrrrrr"
    )
} else {
  # DT::datatable for >20 targets — use raw numeric columns for sorting/filtering
  summary_df %>%
    arrange(pct_covered_1x) %>%
    select(reference_id, ref_length, covered_bases, pct_covered_1x,
           mean_depth, median_depth, pct_covered_2x, pct_covered_5x,
           pct_covered_10x, max_gap_length, num_gaps, pct_near_probe) %>%
    datatable(
      colnames = c("Reference", "Length", "Covered", "%\u22651X", "Mean Depth",
                    "Median Depth", "%\u22652X", "%\u22655X", "%\u226510X",
                    "Max Gap", "Gaps", "% Near Probe"),
      options = list(pageLength = 15, scrollX = TRUE, order = list(list(3, "asc"))),
      rownames = FALSE,
      filter = "top"
    ) %>%
      formatRound(columns = c("mean_depth", "median_depth"), digits = 1) %>%
      formatRound(columns = c("pct_covered_1x", "pct_covered_2x", "pct_covered_5x",
                               "pct_covered_10x", "pct_near_probe"), digits = 1) %>%
      formatCurrency(columns = c("ref_length", "covered_bases", "max_gap_length"),
                     currency = "", digits = 0)
}
```

## Coverage Breadth

```{r coverage-breadth, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 5, fig.width=if (is_large) 10 else 8}
summary_df <- summary_df %>%
  mutate(
    coverage_class = case_when(
      pct_covered_1x >= 95 ~ "Excellent (>=95%)",
      pct_covered_1x >= 80 ~ "Good (>=80%)",
      TRUE ~ "Poor (<80%)"
    ),
    coverage_class = factor(coverage_class,
                            levels = c("Excellent (>=95%)", "Good (>=80%)", "Poor (<80%)"))
  )

quality_colors <- c("Excellent (>=95%)" = "#27ae60",
                     "Good (>=80%)" = "#f39c12",
                     "Poor (<80%)" = "#e74c3c")

if (is_small) {
  # Horizontal bar chart with target names
  ggplot(summary_df, aes(x = reorder(reference_id, pct_covered_1x),
                          y = pct_covered_1x, fill = coverage_class)) +
    geom_col(width = 0.7) +
    geom_hline(yintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.3) +
    coord_flip() +
    scale_fill_manual(values = quality_colors) +
    labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
         title = "Probe Tiling Breadth per Target") +
    theme(legend.position = "bottom")

} else if (is_medium) {
  # Horizontal bar chart, compressed, smaller text
  ggplot(summary_df, aes(x = reorder(reference_id, pct_covered_1x),
                          y = pct_covered_1x, fill = coverage_class)) +
    geom_col(width = 0.9) +
    geom_hline(yintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.3) +
    coord_flip() +
    scale_fill_manual(values = quality_colors) +
    labs(x = NULL, y = "% Bases Covered (>= 1X)", fill = "Quality",
         title = "Probe Tiling Breadth per Target") +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))

} else {
  # Histogram of coverage percentages for large panels
  ggplot(summary_df, aes(x = pct_covered_1x, fill = coverage_class)) +
    geom_histogram(binwidth = 2, boundary = 0, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = quality_colors) +
    geom_vline(xintercept = c(80, 95), linetype = "dashed", color = "grey40", linewidth = 0.5) +
    labs(x = "% Bases Covered (>= 1X)", y = "Number of Targets", fill = "Quality",
         title = "Distribution of Probe Tiling Breadth",
         subtitle = sprintf("%d targets | Dashed lines: 80%% and 95%%", n_targets)) +
    theme(legend.position = "bottom")
}
```

```{r coverage-breadth-table, eval=is_large}
cat("\n\n**Worst-covered targets:**\n\n")
worst <- summary_df %>%
  arrange(pct_covered_1x) %>%
  head(20) %>%
  select(reference_id, ref_length, pct_covered_1x, mean_depth, max_gap_length, num_gaps)

datatable(worst,
  colnames = c("Reference", "Length", "%\u22651X", "Mean Depth", "Max Gap", "Gaps"),
  options = list(pageLength = 10, dom = "tp"),
  rownames = FALSE
) %>%
  formatRound(columns = c("pct_covered_1x", "mean_depth"), digits = 1) %>%
  formatCurrency(columns = c("ref_length", "max_gap_length"), currency = "", digits = 0)
```

## Tiered Coverage

```{r tiered-coverage, fig.height=if (is_small) max(3, n_targets * 0.5) else if (is_medium) 6 else 5}
tiered <- summary_df %>%
  select(reference_id, pct_covered_1x, pct_covered_2x, pct_covered_5x, pct_covered_10x) %>%
  pivot_longer(cols = starts_with("pct_covered_"),
               names_to = "threshold", values_to = "pct") %>%
  mutate(threshold = case_when(
    threshold == "pct_covered_1x"  ~ ">=1X",
    threshold == "pct_covered_2x"  ~ ">=2X",
    threshold == "pct_covered_5x"  ~ ">=5X",
    threshold == "pct_covered_10x" ~ ">=10X"
  ),
  threshold = factor(threshold, levels = c(">=1X", ">=2X", ">=5X", ">=10X")))

tier_colors <- c(">=1X" = "#3498db", ">=2X" = "#2ecc71",
                  ">=5X" = "#f39c12", ">=10X" = "#e74c3c")

if (is_small) {
  ggplot(tiered, aes(x = reorder(reference_id, pct), y = pct, fill = threshold)) +
    geom_col(position = "dodge", width = 0.7) +
    coord_flip() +
    scale_fill_manual(values = tier_colors) +
    labs(x = NULL, y = "% Bases Covered", fill = "Depth Threshold",
         title = "Tiered Probe Coverage per Target") +
    theme(legend.position = "bottom")

} else if (is_medium) {
  ggplot(tiered, aes(x = reorder(reference_id, pct), y = pct, fill = threshold)) +
    geom_col(position = "dodge", width = 0.9) +
    coord_flip() +
    scale_fill_manual(values = tier_colors) +
    labs(x = NULL, y = "% Bases Covered", fill = "Depth Threshold",
         title = "Tiered Probe Coverage per Target") +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))

} else {
  # Box plots for large panels
  ggplot(tiered, aes(x = threshold, y = pct, fill = threshold)) +
    geom_boxplot(alpha = 0.7, outlier.size = 0.5) +
    geom_jitter(width = 0.2, alpha = 0.1, size = 0.3) +
    scale_fill_manual(values = tier_colors) +
    labs(x = "Depth Threshold", y = "% Bases Covered",
         title = "Distribution of Tiered Probe Coverage",
         subtitle = sprintf("%d targets", n_targets)) +
    theme(legend.position = "none")
}
```

## Gap Analysis

```{r gap-analysis, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 4}
gap_data <- summary_df %>% filter(num_gaps > 0)

if (nrow(gap_data) == 0) {
  cat("All targets are fully tiled with no gaps.")
} else if (is_small) {
  # Lollipop chart with labels
  ggplot(gap_data, aes(x = reorder(reference_id, max_gap_length), y = max_gap_length)) +
    geom_segment(aes(xend = reference_id, y = 0, yend = max_gap_length),
                 color = "#e74c3c", linewidth = 1) +
    geom_point(size = 4, color = "#e74c3c") +
    geom_text(aes(label = sprintf("%s bp (%d gaps)",
                                   format(max_gap_length, big.mark = ","), num_gaps)),
              hjust = -0.1, size = 3.5) +
    coord_flip() +
    labs(x = NULL, y = "Largest Gap (bp)",
         title = "Largest Uncovered Stretch per Target",
         subtitle = "Targets with no gaps are omitted") +
    expand_limits(y = max(gap_data$max_gap_length) * 1.3)

} else if (is_medium) {
  # Compact bars, no labels, sorted by gap size
  ggplot(gap_data, aes(x = reorder(reference_id, max_gap_length), y = max_gap_length)) +
    geom_col(fill = "#e74c3c", width = 0.9, alpha = 0.8) +
    coord_flip() +
    labs(x = NULL, y = "Largest Gap (bp)",
         title = "Largest Uncovered Stretch per Target",
         subtitle = sprintf("%d of %d targets have gaps", nrow(gap_data), n_targets)) +
    theme(axis.text.y = element_text(size = 5))

} else {
  # Heatmap-style: one row per target, width = gap size, sorted
  gap_data <- gap_data %>% arrange(desc(max_gap_length)) %>%
    mutate(rank = row_number())

  ggplot(gap_data, aes(x = max_gap_length, y = reorder(factor(rank), -rank))) +
    geom_col(fill = "#e74c3c", alpha = 0.7, width = 1) +
    labs(x = "Largest Gap (bp)", y = NULL,
         title = "Gap Sizes Across All Targets",
         subtitle = sprintf("%d of %d targets have gaps (sorted by gap size, largest at top)",
                            nrow(gap_data), n_targets)) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major.y = element_blank())
}
```

```{r gap-table, eval=(nrow(gap_data) > 0 && !is_small)}
cat("\n\n**Gap details:**\n\n")
gap_table <- gap_data %>%
  arrange(desc(max_gap_length)) %>%
  select(reference_id, ref_length, pct_covered_1x, num_gaps, max_gap_length)

if (nrow(gap_table) <= 20) {
  knitr::kable(gap_table,
    col.names = c("Reference", "Length", "%\u22651X", "Gaps", "Max Gap (bp)"),
    align = "lrrrr"
  )
} else {
  datatable(gap_table,
    colnames = c("Reference", "Length", "%\u22651X", "Gaps", "Max Gap (bp)"),
    options = list(pageLength = 15, order = list(list(4, "desc"))),
    rownames = FALSE,
    filter = "top"
  ) %>%
    formatRound(columns = "pct_covered_1x", digits = 1) %>%
    formatCurrency(columns = c("ref_length", "max_gap_length"), currency = "", digits = 0)
}
```

## Pangenome Probe Depth

Targets concatenated end-to-end, sorted by name.`r if (n_targets <= 12) " Vertical grey lines mark target boundaries." else ""`

```{r pangenome, fig.height=4, fig.width=10}
if (nrow(depth_df) > 0) {
  # Build cumulative offsets for each target
  target_order <- summary_df %>% arrange(reference_id)
  offsets <- target_order %>%
    mutate(offset = cumsum(lag(ref_length, default = 0)))

  # Join offsets and compute pangenome position
  pan_df <- depth_df %>%
    inner_join(offsets %>% select(reference_id, offset), by = "reference_id") %>%
    mutate(pan_position = position + offset)

  # Downsample for rendering speed
  pan_df <- pan_df %>%
    mutate(bin = ceiling(row_number() / max(1, ceiling(n() / 2000)))) %>%
    group_by(bin) %>%
    summarise(pan_position = mean(pan_position), depth = mean(depth),
              reference_id = first(reference_id), .groups = "drop")

  # Target boundary positions for vertical lines
  boundaries <- offsets %>%
    mutate(boundary = offset + ref_length) %>%
    filter(row_number() < n())  # no line after last target

  ggplot(pan_df, aes(x = pan_position, y = depth)) +
    geom_area(fill = "#3498db", alpha = 0.4) +
    geom_line(color = "#2c3e50", linewidth = 0.2) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.3) +
    geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.3) +
    { if (nrow(boundaries) > 0 && n_targets <= 12)
        geom_vline(data = boundaries, aes(xintercept = boundary),
                   color = "grey50", linetype = "solid", linewidth = 0.3) } +
    labs(x = "Pangenome Position", y = "Probe Depth",
         title = "Probe Depth Across Concatenated Targets",
         subtitle = sprintf("%d targets | Dashed lines: 1X (green), 5X (orange)",
                            nrow(target_order))) +
    theme(axis.text.x = element_text(size = 7))
}
```

## Probe Depth Profiles {.tabset}

```{r depth-overview-calc, include=FALSE}
# Determine how many targets to show in faceted overview and detail
n_facet <- if (is_small) n_targets else if (is_medium) min(12, n_targets) else min(12, n_targets)
n_detail <- if (is_small) n_targets else min(10, n_targets)
show_detail <- n_targets <= 100  # skip individual plots for very large panels
```

```{r depth-overview, fig.height=max(3, ceiling(min(n_facet, 12) / 3) * 2.5), fig.width=10}
if (nrow(depth_df) > 0) {
  # Show worst-covered targets in faceted overview
  worst_refs <- summary_df %>%
    arrange(pct_covered_1x) %>%
    head(n_facet) %>%
    pull(reference_id)

  overview_data <- depth_df %>% filter(reference_id %in% worst_refs)

  if (n_targets > n_facet) {
    overview_subtitle <- sprintf("Showing %d worst-covered of %d targets | Dashed lines: 1X (green), 5X (orange)",
                                  n_facet, n_targets)
  } else {
    overview_subtitle <- "Dashed lines: 1X (green), 5X (orange)"
  }

  # Downsample for rendering speed
  plot_data <- overview_data %>%
    group_by(reference_id) %>%
    mutate(
      bin = ceiling(row_number() / max(1, ceiling(n() / 500)))
    ) %>%
    group_by(reference_id, bin) %>%
    summarise(position = mean(position), depth = mean(depth), .groups = "drop")

  ggplot(plot_data, aes(x = position, y = depth)) +
    geom_area(fill = "#3498db", alpha = 0.4) +
    geom_line(color = "#2c3e50", linewidth = 0.3) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.3) +
    geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.3) +
    facet_wrap(~reference_id, scales = "free", ncol = 3) +
    labs(x = "Position", y = "Probe Depth",
         title = "Probe Depth Across Targets",
         subtitle = overview_subtitle) +
    theme(strip.text = element_text(size = 8),
          axis.text = element_text(size = 6))
}
```

```{r depth-detail, results='asis', fig.width=8, fig.height=3, eval=show_detail}
if (nrow(depth_df) > 0 && nrow(summary_df) > 0) {
  ordered <- summary_df %>% arrange(pct_covered_1x)

  if (n_targets > n_detail) {
    cat(sprintf('\n\n*Showing detail for the %d worst-covered targets (of %d total).*\n\n',
                n_detail, n_targets))
  }

  for (i in seq_len(n_detail)) {
    ref_id <- ordered$reference_id[i]
    ref_data <- depth_df %>% filter(reference_id == ref_id)

    if (nrow(ref_data) == 0) next

    s <- ordered[i, ]
    cat(sprintf('\n\n<details><summary><strong>%s</strong> &mdash; %.1f%% covered, %.1fX mean, %s bp, %d gaps (max %s bp)</summary>\n\n',
                ref_id, s$pct_covered_1x, s$mean_depth,
                format(s$ref_length, big.mark = ","),
                s$num_gaps, format(s$max_gap_length, big.mark = ",")))

    # Identify zero-depth positions for gap rug marks
    zero_pos <- ref_data %>% filter(depth == 0)

    p <- ggplot(ref_data, aes(x = position, y = depth)) +
      geom_area(fill = "#3498db", alpha = 0.3) +
      geom_line(color = "#2c3e50", linewidth = 0.4) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "#27ae60", linewidth = 0.5) +
      geom_hline(yintercept = 5, linetype = "dashed", color = "#f39c12", linewidth = 0.5)

    if (nrow(zero_pos) > 0) {
      p <- p + geom_rug(data = zero_pos, aes(x = position), color = "darkgreen",
                        alpha = 0.3, sides = "b", inherit.aes = FALSE)
    }

    p <- p +
      annotate("text", x = Inf, y = 1, label = "1X", hjust = 1.2, vjust = -0.5,
               color = "#27ae60", size = 3) +
      annotate("text", x = Inf, y = 5, label = "5X", hjust = 1.2, vjust = -0.5,
               color = "#f39c12", size = 3) +
      labs(x = "Genome Position", y = "Probe Depth",
           title = sprintf("Probe Depth: %s", ref_id),
           subtitle = sprintf("%.1f%% covered | %.1fX mean | %.1f median | %s bp",
                              s$pct_covered_1x, s$mean_depth, s$median_depth,
                              format(s$ref_length, big.mark = ",")),
           caption = if (nrow(zero_pos) > 0) "Rug: uncovered positions" else NULL) +
      theme_minimal(base_size = 12)

    print(p)
    cat('\n\n</details>\n\n')
  }
}
```

```{r depth-large-notice, results='asis', eval=!show_detail}
cat(sprintf('\n\n*Individual depth plots are omitted for large panels (%d targets). See the faceted overview and pangenome plot above, or examine `probe_depth.tsv` directly.*\n\n', n_targets))
```

## Proximity Coverage

Pull-down zone: `r proximity_bp` bp around each probe.

```{r proximity, fig.height=if (is_small) max(3, n_targets * 0.4) else if (is_medium) 6 else 5}
prox_subtitle_small <- sprintf("Blue: directly covered | Purple: within %d bp pull-down zone", proximity_bp)

if (is_small) {
  ggplot(summary_df, aes(x = reorder(reference_id, pct_near_probe), y = pct_near_probe)) +
    geom_col(fill = "#8e44ad", width = 0.7, alpha = 0.8) +
    geom_col(aes(y = pct_covered_1x), fill = "#3498db", width = 0.7, alpha = 0.5) +
    coord_flip() +
    labs(x = NULL, y = "% Target Bases",
         title = "Direct Coverage vs Proximity Coverage",
         subtitle = prox_subtitle_small) +
    theme(legend.position = "bottom")

} else if (is_medium) {
  ggplot(summary_df, aes(x = reorder(reference_id, pct_near_probe), y = pct_near_probe)) +
    geom_col(fill = "#8e44ad", width = 0.9, alpha = 0.8) +
    geom_col(aes(y = pct_covered_1x), fill = "#3498db", width = 0.9, alpha = 0.5) +
    coord_flip() +
    labs(x = NULL, y = "% Target Bases",
         title = "Direct Coverage vs Proximity Coverage",
         subtitle = prox_subtitle_small) +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 5))

} else {
  # Scatter plot: direct coverage vs proximity coverage
  ggplot(summary_df, aes(x = pct_covered_1x, y = pct_near_probe)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(alpha = 0.5, color = "#8e44ad", size = 1.5) +
    labs(x = "% Directly Covered (>=1X)",
         y = sprintf("%% Within %d bp Pull-Down Zone", proximity_bp),
         title = "Direct Coverage vs Proximity Coverage",
         subtitle = sprintf("%d targets | Diagonal: perfect agreement", n_targets)) +
    coord_fixed(xlim = c(0, 100), ylim = c(0, 100))
}
```

## Multi-Mapping Probes

`r nrow(multi_df)` probes map to more than one target sequence. See `multi_mapping_probes.tsv` for the full list.

```{r multi-mapping, eval=(nrow(multi_df) > 0)}
if (nrow(multi_df) <= 20) {
  multi_df %>%
    arrange(desc(num_targets)) %>%
    knitr::kable(
      col.names = c("Probe ID", "# Targets", "Targets"),
      align = "lrl"
    )
} else {
  multi_df %>%
    arrange(desc(num_targets)) %>%
    datatable(
      colnames = c("Probe ID", "# Targets", "Targets"),
      options = list(pageLength = 15, order = list(list(1, "desc"))),
      rownames = FALSE,
      filter = "top"
    )
}
```

---

*Generated by [BaitBench](https://github.com/CrossLab/BaitBench) probe-coverage*
